### 2019-04-21 main.tex
Initial Overleaf Import

\documentclass{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[numbers]{natbib}
\bibliographystyle{plainnat}
\usepackage{booktabs}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{setspace}
\usepackage{lineno}  % For reviewers
\usepackage{authblk}
\usepackage{blindtext}

% Use pretty font
\usepackage[scaled]{palatino}
\renewcommand\familydefault{\sfdefault} 
\usepackage[T1]{fontenc}

% big letter
\usepackage{lettrine}

% Make a pretty abstract box...
% \renewenvironment{abstract}
% {\begin{quote}
% \noindent \rule{\linewidth}{.5pt}\par{\bfseries \abstractname.}}
% {\medskip\noindent \rule{\linewidth}{.5pt}
% \end{quote}
% }

% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\title{Homeostatic mechanisms may shape the type and duration of oscillatory modulation.}
\author[1,2,*]{Erik J. Peterson}
\author[2,3,4]{Bradley Voytek}

\affil[1]{Department of Psychology. Carnegie Mellon University, Pittsburgh, PA 15213}
\affil[2]{Department of Cognitive Science,~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}
\affil[3]{Neurosciences Graduate Program,~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}
\affil[4]{Hal{\i}c{\i}o\u{g}lu Data Science Institute, University of California, San Diego, 92093~~~~}
\affil[*]{Corresponding author: Erik.Exists@gmail.com~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}
\date{}                     %% if you don't need date to appear
\setcounter{Maxaffil}{0}
\renewcommand\Affilfont{\itshape\small}

\begin{document}
\maketitle
% \linenumbers
% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\begin{abstract}
Neural oscillations are observed ubiquitously in the mammalian brain. However the stability of oscillations is highly variable. Some oscillations are \textit{tonic}, lasting for seconds or even minutes; others are unstable, appearing only as a single-cycle \textit{burst}. In a model of hippocampal neurons, we use numerical simulations to show how these different forms of rhythm stability can interact with activity-dependent homeostasis to profoundly alter the modulatory effect of neural oscillations. Under homeostasis, tonic oscillations that are synaptically excitatory have a paradoxical effect; they decrease excitability and desynchronizing firing. Tonic oscillations that are synaptically inhibitory--like those in a real hippocampus--fail to generate new action potentials and so provoke no homeostatic response. This may explain why the theta rhythm in hippocampus is synaptically inhibitory: inhibitory oscillations don't raise the firing threshold, as excitatory oscillations do, and so can preserve each cell's dynamic range. Based on these simulations, we also speculate that homeostasis may explain why excitatory intra-cortical and intra-layer oscillations often appear as bursts. In our model bursts minimally interact with the slow homeostasis time constant and so retain typical excitatory effects.\\
\end{abstract}

% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\section*{Introduction}
\lettrine[loversize=0,nindent=0,realheight=true]{N}{euromodulation\ and\ homeostasis} are inter-linked but opposing phenomena. Modulation perturbs excitability, which we define as the propensity for a stimulus to elicit an action potential. Homeostasis acts to this quench perturbation, driving the excitability of the cell back to a biologically desirable set point \cite{LeMasson1993,Abbott1993}. While the interplay between chemical modulators and homeostasis has been studied for more than 20 years \cite{LeMasson1993,Abbott1993,Golowasch1999,Marder2014,Marder2015,Gutierrez2013}, the relationship between network-level synaptic modulations--like neural oscillations--and homeostasis is not well understood, on either theoretical or empirical grounds. Like neuromodulators, neural oscillations alter excitability and firing statistics. Uniquely, though, oscillations create synchronous windows of activity \cite{Lisman2013,Voytek2015}. Temporally grouping action potentials into windows improves signal to noise and increases the number of coincident firing events \cite{Chen2013,Zhou2015,Voytek2015a,Peterson2017}, driving learning at individual synapses \cite{Muller2011,Song2000,Markram1997}.

We prepose the same mechanisms that link homeostasis with chemical neuromodulation can also come into play during oscillatory modulation. After all, both kinds of modulation lead to tonic changes in spiking and in Ca\textsuperscript{2+} concentration, which can in turn drives changes to intrinsic homeostasis \cite{Liu1998}.

To test this, we model activity-dependent intrinsic homeostasis in a feed-forward population of hippocampal pyramidal cells \cite{Siegel1994}. In this model homeostasis is mediated by a Ca\textsuperscript{2+}-dependent mechanism \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,OLeary2014} pioneered by LeMasson \cite{LeMasson1993,Abbott1993}. In this model, Ca\textsuperscript{2+} acts a sensor or proxy for tonic changes in the membrane voltage. To counter tonic changes in Ca\textsuperscript{2+} levels, the expression of ion channels is altered, returning the Ca\textsuperscript{2+} level to a predefined ``good'' value \cite{Golowasch1999,OLeary2013}. Following Siegel \cite{Siegel1994}, increases in Ca\textsuperscript{2+} lead to downregulation of Na\textsuperscript{+} and Ca\textsuperscript{2+} channels, and upregulation of fast K\textsuperscript{+} channels. To minimize the effect of homeostasis on the rise and fall of action potential dynamics, we also add a KCa channel which is not present in Siegel \citep{Siegel1994}. In real cells intrinsic homeostasis changes the expression level, membrane trafficking, and kinetics of ion channels. Here, these details are not directly simulated. Instead we mimic the net or bulk effect of all these changes by altering the maximum conductance of individual ionic channels, which are modelled using Hodgkin-Huxley terms \cite{LeMasson1993,OLeary2013,OLeary2014}.

Given that chemical modulators operate on both short and long times scales \cite{Marinelli2014,Marder2014,Cohen2015,Daw2002}, we examined two timescales of oscillation. We contrast the effects of shorts bursts of oscillation to long lasting tonic rhythms, both of which are observed {\textit{in vivo}}, and which may play different physiological, cognitive, and computational roles \cite{Lundqvist2016,vanEde2018}. We also explore how oscillation duration interacts with synapse type, examining both AMPA- or GABA-ergic oscillations.

In our highly simplified model of hippocampal cells, intrinsic homeostasis can profoundly alter the modulatory effect of neural oscillations. \textit{Tonic} excitatory modulation, paradoxically, generates a homeostatic response that increases the firing threshold. This suppresses excitability which, in turn, desynchronizes population activity. \textit{Bursts} of excitation, meanwhile, don't show homeostatic suppression, and may even benefit from a weak level of homeostasis. This is due the constant noisy background level input also present in the model, which homeostatic mechanisms help control. Inhibitory GABAergic oscillations, however, show little to no homeostatic effect, suggesting that inhibitory oscillations might better isolate any phase coding scheme from the stimulus-driven response. This is might be important, for example, in hippocampal phase-coding schemes of memory \cite{Lisman2013}.

% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\section*{Results}
We study Ca\textsuperscript{2+}-mediated homeostasis in a feed-forward network, using single compartment neurons. We modulate this network using neural oscillations. Oscillations are simulated either as tonic, lasting for the entire experiment, or may as a burst at the end, where it overlaps with input stimulus (Figure~\ref{fig:f1}\textbf{a}). Input into the model has both a constant background level of noise (not shown), and a stronger stimulus that is delivered once homeostatic equilibrium is reached \cite{Barth2012} (Figure~\ref{fig:f1}). During this strong stimulus we measure the network's response and how population firing and synchrony are affected by homeostasis under different oscillatory regimes.

Each experiment began with an instantiation of a randomly generated feed-forward network. A single neuron from this network is depicted in Figure~\ref{fig:f1}\textbf{b}. This network is subjected to a range of modulatory and control conditions, including oscillatory strength, duration, and synapse type (AMPA or GABA). Each experiment lasted 20 seconds. In Figure~\ref{fig:f1}\textbf{c} we depict key aspects of model output during an experiment. 
In real systems, intrinsic homeostasis is thought to happen over minutes or days. However, simulation times that are hours or days long are not computationally feasible. As a result, we follow the field and study a model where Ca\textsuperscript{2+} dynamics happen with a 4-second half-life, denoted by $\tau_h$. Despite the radical difference between real and simulated time-scales, all that matters mathematically is that Ca\textsuperscript{2+} dynamics happen much slower than all the other synaptic/membrane dynamics. In practice, this means a timescale of $\tau_h > 4$ seconds is a reasonable first-order approximation \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,Marder2014,OLeary2014,LeMasson1993,Abbott1993}.

After homeostatic equilibrium is reached, we measure two features: the synchrony between action potentials (measured by the Kappa correlation) and changes in the excitability of the system (measured as a change in population firing rate). Both of which are defined in the \textit{Methods}. To ensure a consistent comparison between experiments, measurements were made over the same 4-cycle or 0.5 second period in all simulations.

% At the end of each experiment we measure three main features: synchrony between action potentials between different neurons, for a given type of modulation; changes in the excitability of the system--measured as a change in population firing rate--and; the ``computational error'' introduced by the combination of modulation and homeostasis. Error is defined as the mean squared error (MSE) measured by comparing spike-times recorded during experimental conditions to recordings made in identical populations simulated without modulation/homeostasis. To ensure a consistent comparison between experiments, measurements were made over the same 4-cycle period \emph{for all simulations}.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{fig1.png}
\caption{\label{fig:f1}
    Diagram of the model. \textbf{a}. Inputs into the model. Top panel depicts the stimulus, applied over the last 0.5 seconds of each simulated trial. Middle and bottom panels depict the two modes of oscillation we examined--tonic and bursting. 
    \textbf{b}. Illustration of a single model neuron, its major currents, its inputs (top arrow) and its output (bottom arrow).
    \textbf{c.} Examples of model output, including membrane potential (top panel), Ca\textsuperscript{2+} concentration, both observed (solid) and the homeostatic target (dotted line). The bottom panel in c. shows homeostatic dynamics beginning with trial onset, followed by a delay to equilibrium over 20 seconds. Note that there is no homeostatic response to stimulus onset at 19.5 seconds. Also note the log scale on the y axis in this panel.
}
\end{figure}

% -----------------------------------------------------------------------
\subsection*{Excitatory modulation.}
Homeostasis completely inverts the effect of tonic AMPAergic oscillatory modulation. To see how, first consider how the network responds to a tonic excitatory modulation without homeostasis. Without homeostasis, increasing the strength of the excitatory oscillation increases excitability, leading to an increase in the population firing rate. Additionally, the neuromodulatory effect of the oscillations is such that action potentials are grouped by oscillation phase, resulting in an increase in population synchrony (Figure \ref{fig:f2}\textbf{b}, black line). With homeostasis this pattern inverts. As excitatory oscillatory modulation strength increases, synchrony and excitability \emph{decreases} (Figure \ref{fig:f2}\textbf{c-d}). Homeostatic mechanisms in the model cause what \emph{should be} excitatory synchronizing modulation to become suppressive and desynchronizing. The stronger the oscillation, the more suppressive the result. By the time the oscillation is about half the strength of the stimulus (which we fix at a firing rate of 6 Hz) the stimulus is completely suppressed (\textit{i.e.}, the population firing rate approaches 0) (Figure \ref{fig:f2}\textbf{c}).

We compared the effect of tonic oscillation to short 4-cycle bursts of excitatory modulation, presented only during the stimulus. Here, the oscillation period is far too short to engage any additional homeostasis. This means that increases in oscillatory strength continue to increase population firing rate and synchrony (Figure~\ref{fig:f2}\textbf{e} and \textbf{f}).

Our model suggests tonic oscillations can profoundly alter coding properties of synaptically excitatory oscillations. This means that bursts of excitatory oscillation are \textit{qualitatively distinct} from their tonic counterparts. The decrease in excitability caused by tonic oscillation can be explained by a direct homeostatic change. Larger tonic oscillations lead to larger tonic increases in the membrane potential, which in turn raise Ca\textsuperscript{2+} levels. The homeostatic equations respond to this change Ca\textsuperscript{2+} by decreasing the conductance of the Na and Ca channels, and increasing the conductance of K and KCa channels (Eq~\ref{eq:dgdt}). The net effect of these dynamics is a increase in the firing threshold.

The inversion in Kappa seen in Figure~\ref{fig:f2}\textbf{d} as well as in \ref{fig:f4}\textbf{b}, is an artifact of the decrease in excitability; it's a low effect $N$ effect. As population spiking becomes more strongly suppressed the total number of spikes declines to the point where the bins used to calculate Kappa often contain no spikes. This in turn inflates Kappa values.

\subsection*{Inhibitory modulation.}
Both sustained tonic and bursting inhibitory oscillatory modulation do not lead to a homeostatic response in our model. In all cases, as oscillatory strength increases, population firing declines dramatically (grey lines in Figure \ref{fig:f2}\textbf{a-d} and light yellow lines in panel \textbf{e-f}).

% \subsection*{The synaptic nature of hippocampal theta.}

% \subsection*{Modulation, homeostasis, and error.}
% When increasing or decreasing the firing rate of a population, or its synchrony, action potentials can be added or re-arranged arbitrarily. We've suggested previously that an efficient modulator is one that adds or moves spikes as little as possible, while achieving a set level of effect \cite{Peterson2018a}. We define spike-time deviations introduced by modulation as an error compared to the original, unmodulated, population output (see Eq~\ref{eq:error} and \cite{Peterson2018a} for more background). We compare observed errors with and without homeostasis, for burst and tonic excitatory modulation. Given that synaptic inhibitory oscillatory modulation failed to trigger any homeostatic effects we don't include it in these analyses.

% Under homeostasis, tonic excitatory oscillations suppress the network responses to stimulation, leading to extremely high errors compared to an identical model simulated without homeostasis included (Figure~\ref{fig:f3}\textbf{a}). Bursts of excitatory oscillatory modulation, however, actually \textit{reduce} error compared to their non-homeostasis counterpart (Figure~\ref{fig:f3}\textbf{a}.). This decrease in error is driven by the baseline level of noise present in our stimulus model, which when input into the hippocampal population leads to a slight reduction in the excitability, this then positively interacts with the modulatory burst which ends more effective modulator compared to the baseline--homeostasis free--model.

\begin{figure}
\centering
\includegraphics[width=1\textwidth]{fig2.png}
\caption{\label{fig:f2}
    The effect of oscillatory modulation on synchrony--which is measured using the Kappa correlation (Eq~\ref{eq:kappa})--and excitability, measured by the population firing rate. Tonic oscillations are shown in grey and black. Bursts are shown in light and dark yellow.
    \textbf{a.}-\textbf{b.} Increases in tonic modulation strength, without homeostasis. This is our reference condition. Top panel (a.) is the observed population firing rate averaged over the 0.5 second stimulus. Bottom (b.) is synchrony over the same period.
    \textbf{c.}-\textbf{d.} Same experiment as a-b but with Calcium-mediated homeostasis, showing how homeostasis with tonic AMPA oscillations reduces population firing and synchrony.
    \textbf{e.}-\textbf{f.} Burst modulation, presented during the stimulus period (4 cycles of oscillation, onset time: 19.5 s).
    \textbf{g.} Change in excitability between bursts and tonic rhythms for all oscillation firing rates. Asterisks denote a significant difference using the Wilcoxon rank sum test ($W = 1886.5, p < 2.2e-16$). The frequency of the oscillatory rhythm was fixed at $f = 8$ in all models.}
\end{figure}

% \begin{figure}
% \centering
% \includegraphics[width=0.6\textwidth]{fig3.png}
% \caption{\label{fig:f3}
%     Synchrony versus error during excitatory modulation, with (black) and without (grey) homeostasis.
%     \textbf{a.} Tonic modulation.
%     \textbf{b.} 4-cycle burst.
% }
% \end{figure}

% -----------------------------------------------------------------------
\subsection*{The effect of Ca\textsuperscript{2+} concentration.}
When the model is run with only stimulus-driven homeostasis, the Ca\textsuperscript{2+} concentration equilibrates to about 0.003 mM. We used this as a standard target value for all modulation experiments, until now. When we vary this value in 0.0002 mM increments, population rate and synchrony either increases or decreases depending on whether the Ca\textsuperscript{2+} increases or decreases, shown Figure \ref{fig:f4}\textbf{a-b}. However despite different initial Ca\textsuperscript{2+} concentrations, each model still shows an identical set of trends as the strength of the oscillation increases (Figure \ref{fig:f4}). That is, increasing or decreasing the target concentration shifts the overall excitability of the network, in an approximately linear way. This means that while the initial choice of 0.003 mM was arbitrary, the qualitative pattern of results we report is not dependent on this choice.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{fig4.png}
\caption{\label{fig:f4}
    Synchrony and excitability shift linearly with the target Calcium concentration, [Ca]. Plotted here is the effect of Ca\textsuperscript{2+} on a tonic excitatory modulation. The black value (0.003 mM) is the standard reference value used in all previous simulations.
    \textbf{a.} Change in population hippocampal population rate for different levels of target [Ca] (colors) as a function oscillation strength (Osc. rate on the x-axis). All values are referenced to a no-modulation control.
    \textbf{b.} Change in population synchrony for different levels of target [Ca] (colors) as a function oscillation strength (firing rate). All values are referenced to a no-modulation control.
    \textbf{c.} Change in population rate as the oscillation duration approaches a more realistic $\tau_h$, the half-life of the homeostasis dynamics (Eq.~\ref{eq:dgdt}). In this control experiment we used a more biologically realistic $\tau_h$ of 600 s (or 10 minutes). All other simulations in the report use a $\tau_h$ of 4 seconds, which is well below most reports of this value in real systems. However in choosing such as small value we follow the majority of the homeostasis modeling literature (for more on this see the \textit{Discussion}).
}
\end{figure}

% -----------------------------------------------------------------------
% -----------------------------------------------------------------------
% -----------------------------------------------------------------------
\section*{Discussion}
Our scientific understanding of homeostasis has been shaped as much by theoretical work as empirical \cite{Marder2014}. In an attempt to understand the interaction between oscillatory modulations and homeostasis, we begin by studying one of the simplest models used in early studies of homeostasis--a population of point neurons \cite{LeMasson1993}.

The feed-forward model of hippocampal pyramidal cells studied here serves as a simple initial model to answer three basic questions. One, do tonic oscillations--generated with biologically-consistent parameters--engage homeostatic mechanisms? Two, does homeostasis in turn change the oscillation's function? Three, do short bursts of oscillation have distinct effects from tonic oscillations? Put another way: can homeostasis help explain why some oscillations, such as hippocampal theta, tend to appear as tonic rhythms while other oscillations tend to appear as bursts? Our results suggest that homeostasis can explain why hippocampal theta is synaptically inhibitory, and why cortical oscillations often appear as bursts.

Bursting, rather than sustained, oscillations tend to be common in the cortex. One striking example is in motor cortical regions, where beta (~12-30 Hz) bursts are prevalent and likely functional. Specifically, beta bursts are very short--sometimes lasting only one or two cycles \cite{Sherman2016}--and relate to self-timed movements \cite{Feingold2015}. Patients with Parkinson's disease show increased motor rigidity and bradykinesia, symptoms associated with prolongation of beta bursts \cite{Tinkhauser2017}. Levodopa treatment was shown to decrease burst probability and duration, and that decrease in burst duration correlated with motor improvement \cite{Tinkhauser2017}.

While many neocortical oscillations tend to be bursty, a counterexample is visual cortical alpha, which can become tonic and high power when humans rest with their eyes closed. Even though both cortical and sub-cortical alpha generators are synaptically excitatory, this rhythm had been classically and paradoxically associated suppression of excitability \cite{Jensen2002,Bonnefond2012,Peterson2017}. While several competing explanations have been offered \cite{Bonnefond2012,Lange2013,Peterson2017} for this paradox, our work raises another possibility. Though we modelled hippocampal cells, the same Ca\textsuperscript{2+} homeostatic mechanisms exist in neocortex. Meaning strong and tonic alpha oscillations, combined with homeostasis, directly suppress population firing in visual cortex. Such an effect, were it to occur, would last well past the moment of oscillation offset. In fact, such long term effects of alpha have been reported in the literature, though the physiological mechanism was often unclear. Our work suggests that intrinsic homeostasis may underlie these effects.

\subsection*{Limitations}
The nature of our model--that fact we use point neurons with only 6 currents--or the fact that our model is strictly feed-forward--without lateral or recurrent connections--means we don't know with confidence to what degree our model's effects will appear in more complex models, or in real neural systems. For example, the primate or rat hippocampus. We do know, however, that oscillations are a ubiquitous feature of cortical activity, as is Ca\textsuperscript{2+}-mediated intrinsic homeostasis. This means the ingredients for oscillation and homeostasis to interact are omnipresent in both sub-cortical and cortical areas.

Homeostatic interactions depend on a number of factors specific to each cell and circuit. These include the duty cycle, power, and frequency of an oscillation, as well as on synaptic strengths and their location in the dendritic tree (an idea we return to below). It also depends on the other inputs into the cell, both from fast synaptic transmission and other (slower) modulators, as well a connection type; simulation studies suggest that recurrent connections can strongly interact with homeostatic regulation \cite{Harnack2015}. The temporal and spatial scales of these factors will strongly influence Ca\textsuperscript{2+} dynamics, which is in turn central to governing what, if any, homeostatic effects oscillations may generate.

Synaptic homeostasis may also play a role in tuning a neuron's response to all types of oscillatory perturbations, as both excitatory and inhibitory synapses are susceptible \cite{Cannon2016}, though in the case of excitatory modulation synaptic and intrinsic homeostasis appear to be linked \cite{Joseph2017}. Still, the exact nature of the response depends on how, and to what degree, oscillatory input and other sensory or internally driven ``computational'' inputs share synapses. This in turn requires considering complex dendritic arbors and their effect on homeostasis \cite{LeMasson1993}, neuromodulation \cite{Jadi2012,Jadi2014}, and computation \cite{Mainen1996,Polsky2004,Mel2004}. Considering these interacting effects together is the next step needed to develop a clearer biological understanding of modulatory oscillations. For inhibitory oscillations the case is even more complex: Ca\textsuperscript{2+} in these cells does not appear to regulate intrinsic homeostasis, but synaptic homeostasis is under a separate mechanism of control \cite{Joseph2017}.

\subsection*{Previous work}
Homeostasis has been extensively studied in the rhythmic pacemaker present in the crab stomatogastric ganglion. Here homeostasis has been shown to stabilize self-organized oscillations \cite{Golowasch1999}, and interact with neuromodulation in a \textit{highly} state dependent way \cite{Marder2014,Marder2015,Marder2014}. However, this work has focused on how chemical neuromodulators affect the formation of a pacemaker, not how a pacemaker can modulate another, downstream, circuit. Which is our focus here.

The interaction between homeostasis and oscillations has previously been considered when the oscillatory input is treated as a signal, not a modulator. Cannon and Miller \citep{Cannon2017} explored how synaptic homeostasis can effectively minimize the effect of modulatory perturbations, thus maximizing mutual information between an incoming oscillatory signal and a single cell's firing pattern. Our analysis could be considered an inverse complement to \citep{Cannon2017}--we study how to minimize the perturbation caused by a modulatory oscillator, rather than how to maximize the transmission of an oscillator.

\subsection*{Conclusion}
 Here, using a relatively simple model of hippocampal neurons, we observe a surprising--even paradoxical--result: that homeostatic effects can invert a normally synchronizing excitatory oscillatory neuromodulator and cause it to become inhibitory and desynchronizing.

 Based on our simple model, we conjecture that intrinsic homeostasis may explain why tonic theta rhythms in the hippocampus are synaptically inhibitory. To make this clear, consider the alternative. If the theta rhythm was strong, tonic, and synaptically excitatory, our model suggests this could lead to an equally strong--but opposing--homeostatic response. According to our model, such as response means that the firing threshold increases and the likelihood a hippocampal neuron can respond to any given stimulation would decrease, perhaps markedly so.

In effect, strong excitatory oscillations consume a substantial portion of each cell's possible dynamic range. On the other hand inhibitory oscillations do not generate an intrinsic homeostatic response, and so leave the dynamic range of the neurons intact. This may also explain why neocortical oscillations tend to be short and bursty, and why some neurological disorders, such as Parkinson's disease, are associated with prolonged rhythms.

% Computational modeling plays a critical role in neuroscience. Explicit computational modeling moves the field from "word model" theories to self-consistent mathematical formalizations [REF]. While these models can be wrong, they offer testable novel venues for research that are often difficult to find due to the immense complexity of the rich, dynamic neuronal environment.

\section*{Methods}
\subsection*{Mathematical model}
We model a feedforward network of hippocampal neurons, subjected to oscillatory modulation. This is instantiated as $N = 2000$ input cells connected to $M = 100$ Hodgkin-Huxley neurons. The $M$ cells in the network were tuned to mimic regular firing \cite{Borgers2005,Borgers2008}. The firing pattern of each input cell (both stimulus and modulation) is from a Poisson process, with a time-varying rate. $N_o$ cells oscillate. $N_s$ serve as input. For simplicity, we let $N_o = \frac{N}{2}$ so $N_o = N_s$. All input cells have a $p = 0.1$ connection probability to the hippocampal population. The synaptic weights for all $N \rightarrow M$ connections $w$ were independently sampled from a uniform distribution, $w \sim \mathcal{U}(5, 50)$ $\mu$S. The firing rate of the oscillating population was governed by sinusoidal pacemaker, with amplitude $A$ and frequency $f$, with the exact form $r_o / 2 (1 + \text{sin}(2 \pi f t)$. As a result, $r_o$ defines the peak firing of the oscillation. The stimulus population was simply modeled by a fixed rate of 6 Hz ($r_s$). The background firing rate $r_b$ was constant, and set at 2 Hz.

Hodgkin-Huxley dynamics were governed by 4 active ionic currents ($I_{Na}$, $I_{K}$, $I_{KCa}$, $I_{Ca}$) and the passive leak current ($I_l = g_l (E_l - V)$). Besides $I_{Ca}$ (which is discussed below), active currents are governed by the standard Hodgkin and Huxley form \cite{Hodgkin1952}. Where $m$ and $h$ respectively track the opening and closing channel kinetics, and $E$ is the channel appropriate Nernst reversal potential. See \textit{Table 1} for the complete set of parameters.

\begin{equation}
\label{eq:Idef}
I = \bar{g} m^p h^q (E - V)
\end{equation}

The Ca\textsuperscript{2+} current $I_{Ca}$ was governed by a form taken from the Morris-Lecar model \cite{Morris1981,LeMasson1993,Siegel1994}.

\begin{equation}
I_\text{Ca} = g_\text{Ca} [1 + \text{tanh} \Big ( \frac{V_1 - V}{ V_2} \Big ) (V_\text{Ca} - V)]
\end{equation}

Overall membrane dynamics were governed by these internal ion conductances, a variable bias current $I_\text{bias}$, and the excitatory synaptic input term $I_S$, which contains both background, stimulus, and oscillatory terms. All synaptic input was in turn governed a single exponential kinetics, which we denote generically using an $x$ subscript below. Though each synaptic input had different inputs, all models shared the same parameters. That is, $\tau_x = \tau_b = \tau_s = \tau_o$ and $\bar{g_x} = \bar{g_b} = \bar{g_s} = \bar{g_o}$.

\begin{equation}
    I_S = -g_b (E_s - V) + -g_s (E_s - V) + -g_o (E_o - V) 
\end{equation}
\begin{equation}
    \tau_x \frac{dg_x}{dt} = -g_x + \bar{g_x} \delta(t - t_x^j)
\end{equation}

\begin{equation}
C \frac{dV}{dt} = I_l + I_{Na} + I_{K} + I_{KCa} + I_{Ca} + I_{\text{bias}} + I_S
\end{equation}

The intrinsic excitability is regulated by altering both inward and outward conductances in response to changes in Ca\textsuperscript{2+} concentration. Following the previous work \cite{LeMasson1993} and \cite{Siegel1994}, we modeled this by allowing the maximal conductances $\bar{g_\text{Na}}$, $\bar{g_\text{K}}$, $\bar{g_\text{KCa}}$, and $\bar{g_\text{Ca}}$ to non-linearly vary in response to changes in Ca\textsuperscript{2+} concentration, $\text{Ca}$. During homeostatic equilibration, conductances drifted until the target Ca\textsuperscript{2+} concentration was met, denoted as $C_T$. In a control experiment a range of $C_T$ values were explored (Figure \ref{fig:f4}), though simulations defaulting to 0.03 mM; the value the system reaches with stimulation ($r_s = 6$ Hz) without modulation ($r_o = 0$). The $\pm$ symbol in equation \ref{eq:dgdt} denotes the direction of ion flow and is $(+)$ for inward going currents (Na and Ca) and $(-)$ for outward going Potassium.

\begin{equation}
\label{eq:dgdt}
\tau_h \frac{d \bar{g_x}}{dt} = \frac{G_x}{1 + e^{\pm (Ca - C_T)/\Delta}}
\end{equation}

Ca\textsuperscript{2+} dynamics were assumed to follow first order kinetics, driven by the Ca\textsuperscript{2+} influx current and clearance rate constant $k$. Values for both $\gamma$ and $k$ were taken from \cite{Liu1998}.

\begin{equation}
\frac{d\text{Ca}}{dt} = -k \text{Ca} - \gamma I_\text{Ca}
\end{equation}

% \subsection*{Estimating excitability, synchrony, and error}
\subsection*{Estimating excitability and synchrony}
We measure excitability by comparing average firing rate of all $M$ neurons in an experiment, with ($r_m$) and without modulation ($\bar{r}_m$). This accounts for any homeostatic drift driven only by the background noise:

\begin{align}
    \label{eq:excite}
    \Delta r = \bar{r}_m - r_m
\end{align}

We measure synchrony using $\kappa$, a binned measure of spiking covariance \cite{Wang1996}. Where $X(l) = 0 \ \text{or} \ 1$ and $Y(l) = 0 \ \text{or} \ 1$ for $l = \{1, 2, ..., K \}$ and with $T/K = \tau$.

\begin{align}
    \label{eq:kappa}
    \kappa_{ij}(\tau) = \frac{\sum_l^K X(l)Y(l)}{\sqrt{\sum_l^K X(l) \sum_l^k Y(l)}}
\end{align}

\begin{align}
\kappa(\tau) = \frac{1}{N} \sum_i^N \sum_j^N \kappa(\tau)_{ij}
\end{align}

% Modulation is optional. Each of the different modulatory systems--oscillations, dopamine, and acetylcholine--to name a few among dozens of options, may be absent, weakly present, or operating at high level at any given moment of time. We therefore reason it is important to quantify how coding is perturbed under varying levels of modulation and homeostasis. We estimate this by comparing differences in spike timing between an experimental condition and a control run with without modulation.

% Deviations in spike-timing caused by modulation are treated as errors, which we quantify with the mean squared error (Eq.~\ref{eq:error}). The more efficient a modulation, the less error it should introduce for a given change in either excitability or synchrony.

% \begin{equation}
% \label{eq:error}
% \text{MSE} = \frac{1}{K} \sum_{i=1}^k (\hat{y} - y_i)^2
% \end{equation}

% In its mathematical form the MSE assumes the reference $\hat{t}$ and observed variables $y$ to be the same length. Here this constraint could not be satisfied: as $r_o$ increases, new action potentials are inevitable. To accommodate potentially ragged lengths we truncated the longer of the two series to match the shorter. For example, if the target series has length $K$, and the observed sequence has length $L$, when $K > L$ then $K-L$ elements are removed, starting with the largest value. When $L > K$, $L - K$ of the largest were eliminated.

% -------------------------------------
\newpage
\begin{table}
\centering
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Symbol} & \textbf{Range (unit)} & \textbf{Description} \\ \midrule
$f$ & 8 (Hz) & Oscillation frequency \\
$r_o$ & 0 - 6 (Hz) & Oscillation firing rate \\
$r_s$ & 6 (Hz) & Stimulus firing rate \\
$C$ & 1 ($\mu$F) & Membrane capacitance \\
$\tau_h$ & \textgreater 4 (sec) & Homeostasis time constant \\
$w$ & 5 - 50 ($\mu$S) & Synaptic weight \\
$\tau_e$ & 5 (msec) & Excitatory synaptic time constant \\
$V_e$ & 0 (mV) & Excitatory synaptic reversal potential \\
$p$ & 0.1 & Connection probability \\
$C_T$ & 0.0028-0.0032 (mM) & Target Ca\textsuperscript{2+} concentration \\
$G_\text{Na}$ & 360 ($\mu$S) & Initial Na conductance \\
$G_\text{K}$ & 120 ($\mu$S) & Initial K conductance \\
$G_\text{KCa}$ & 60 ($\mu$S) & KCa conductance \\
$g_\text{Na}$ & 180 ($\mu$S) & Max. Na conductance \\
$g_\text{K}$ & 60 ($\mu$S) & Max. K conductance \\
$g_\text{KCa}$ & 30 ($\mu$S) & Max. K conductance \\
$g_\text{Ca}$ & 0.03 ($\mu$S) & Max. Ca\textsuperscript{2+} conductance \\
$g_l$ & 1 ($\mu$S) & Leak conductance \\
$V_l$ & -70 (mV) & Leak reversal potential \\
$V_K$ & -100 (mV) & K reversal potential \\
$V_\text{Na}$ & 50 (mV) & Na reversal potential \\
$V_\text{Ca}$ & 150 (mV) & Ca\textsuperscript{2+} reversal potential \\
$V_\text{1}$ & -50 (mV) & Morris-Lecar constant \\
$V_\text{2}$ & 10 (mV) & Morris-Lecar constant \\
$\Delta$ & 0.6 ($\mu$M) & Ca\textsuperscript{2+} influx rate \\
$k$ & 1/200 (1 / msec) & Ca\textsuperscript{2+} clearance rate \\
$\gamma$ & -0.00047 (mM / mA / msec) & Ca\textsuperscript{2+} current conversion constant \\
$\delta t$ & 0.01 (msec) & Integration time step \\ \bottomrule
\end{tabular}
\caption{Model parameters.}
\label{table:params}
\end{table}

% -----------------------------------------------
\newpage
\bibliography{library_min}

% -----------------------------------------------
\end{document}

### 2019-04-22 main.tex
Updates from Overleaf

@@
 \usepackage{blindtext}
+\usepackage{fancyhdr}
 
@@
 
-% Make a pretty abstract box...
-% \renewenvironment{abstract}
-% {\begin{quote}
-% \noindent \rule{\linewidth}{.5pt}\par{\bfseries \abstractname.}}
-% {\medskip\noindent \rule{\linewidth}{.5pt}
-% \end{quote}
-% }
+\pagestyle{fancy}
+\rhead{Homeostasis and oscillations}
 
@@
 \maketitle
-% \linenumbers
+\linenumbers
 % -----------------------------------------------------
@@
 % -----------------------------------------------------
+\noindent
+\textsc{\textbf{New and Noteworthy}} In a model of hippocampal neurons we report a paradoxical result: Ca\textsuperscript{2+}-mediated homeostasis causes normally synchronizing AMPAergic oscillations to become inhibitory and desynchronizing.
+We conjecture this could be why the theta rhythm is inhibitory.
+In pyramidal cells GABA doesn't generate new action potentials and so provokes no homeostatic response.
+The intricate interplay of neuromodulators, like acetylcholine, with homeostasis is well known.
+The interplay between oscillations and homeostasis might be as complex, and as important.
+
 \section*{Introduction}
-\lettrine[loversize=0,nindent=0,realheight=true]{N}{euromodulation\ and\ homeostasis} are inter-linked but opposing phenomena.
+\lettrine[loversize=0,nindent=0,realheight=true]{N}{} euromodulation and homeostasis are inter-linked but opposing phenomena.
 Modulation perturbs excitability, which we define as the propensity for a stimulus to elicit an action potential.

### 2019-04-22 main.tex
Updates from Overleaf
Added missing source code url

@@
 
-We prepose the same mechanisms that link homeostasis with chemical neuromodulation can also come into play during oscillatory modulation.
+We propose the same mechanisms that link homeostasis with chemical neuromodulation can also come into play during oscillatory modulation.
 After all, both kinds of modulation lead to tonic changes in spiking and in Ca\textsuperscript{2+} concentration, which can in turn drives changes to intrinsic homeostasis \cite{Liu1998}.
@@
 To ensure a consistent comparison between experiments, measurements were made over the same 4-cycle or 0.5 second period in all simulations.
+Source code for all simulations is available at \url{https://github.com/voytekresearch/resistingrhythm}
 

### 2020-03-06 response_JNP_1.md
Reviewer response 1 (JNP) complete

# Response 1

# Reviewer #1 (Comments to the Author):
# ------------------------------------

In this manuscript, the authors developed a simplified computational model able to explore basic interactions between neural oscillations and homeostasis. They found that oscillations that differ in the time domain (tonic or burst of oscillations) also affect differently to the homeostasis of the system, in a non trivial way, what could be directly related to some of the proposed underlying mechanisms of neural information processing during cognitive functions. I find the idea of the manuscript very interesting and novel, the proposed questions are properly addressed and limitations are clarified. The main finding that tonic excitatory oscillations desynchronize firing under homeostatic changes is of great interest and open new interesting interpretations of several experimental phenomena available in the field. I would recommend to clarify a few minor points before publication:

1) The authors adressed the impact of homeostasis in tonic and burst-like oscillations, both of them considered at 8Hz. What would be the impact of their model when considering oscillations in different ranges of the frequency domain? The model is based in a feed-forward population of hippocampal pyramidal cells and it is known that in the hippocampus both types of oscillations (tonic and burst-like) are present. An example of tonic would be theta oscillations (8 Hz), which underlying inhibitory drive the authors interpreted as possibly related with their results. On the other hand, burst-like oscillations exist in the hippocampus such as different gamma band oscillations (60 Hz, 90 Hz) and ripples (150 Hz). In addition to cortical oscillations, if the authors could interpret their findings in terms of these different burst-like hippocampal oscillations, it would enhance the message of the paper since it would be easier to relate to empirical data.

> This is very nice point. We did three things in response. Most important we added simulations of both 60 and 90 Hz gamma, and included these in a new Figure 3. As expected, suppression happens in gamma as well. We also, for the sake of completeness, studied small variations in theta. Real theta frequencies/speeds drift during behavior. Again and as expected small drifts don't seem to have any effect on suppression (Figure 5d-e). We also clarified again that real theta is GABAergic, and AMPA simulations were done only to show the suppression effect in controlled and consistent way.

> The new text reads,

> Of course theta oscillations in hippocampus are GABAergic, not AMPAergic as we model here. Exploring this artificial oscillation lets us show the central principle we want to highlight in our simple model. It seems that tonic oscillations--when joined with homeostasis--can profoundly alter the coding properties of what one would expect from oscillatory modulation. The degree of effect depends on the degree to which the oscillation changes the cell's tonic Ca$^{2+}$. It also shows that because of homeostasis, bursts are \textit{qualitatively distinct} from their tonic counterparts.

> The specific effect we observed (suppression of excitability) is seen because the homeostatic equations that \textit{we chose} respond to changes in Ca\textsuperscript{2+} by decreasing the conductance of the Na and Ca channels, and increasing the conductance of K and KCa channels (Eq~\ref{eq:dgdt}). The net effect of these dynamics is to decrease the sodium channel conductance, and increase the potassium. This makes firing an action potential less likely, for any given input. These channel dynamics are visualized in Figure~\ref{fig:f1}\textbf{c} and the corresponding FI curve is shown in Figure~\ref{fig:f4}.

> The general effect we would expect for some other selection of channels, chosen by some other theorists or by evolution itself, depends on those channels. O'Leary \cite{OLeary2014} has shown how complex the relationship is between homeostasis, channel conductance, and neuronal firing patterns. It's complex enough that general statements are probably impossible. One one hand, our simple model has been shown by others to recapitulate hippocampal pyramidal cell's activity in laboratory settings \cite{LeMasson1993}. On the other, it is \textit{very simple} and so most probably incomplete at best. Experimental testing is needed.

2) In Figure 2 c.d, the model predicts a decrease of both synchrony and excitability for AMPA driven tonic oscillations with increasing oscillation strength. However, due to the low excitability that the system reaches around half of the strength of the oscillation, an inflection point and succesive discontinuity appears for the syncrhony. The authors explain this in the paper as a result of the mathematical limitations of the model. However, physiologically, it is mostly not feasible that the system reaches an excitability point where the population rate is constant zero for long periods of time. Do the authors think that modifying the implementation of the background level noise in their model this could be corrected?

> In response we halved the oscillation strength in all simulations. It now peaks at firing rate of 3 Hz. It was 6 Hz before. This largely removed this unrealistic model artifact. It also (mostly) prevented an unrealistic inversion of synchrony seen in some of the control experiments.

After these points have been addressed, I highly recommend the publication of this very interesting study this Journal.

> Thank you!

# Reviewer #2 (Comments to the Author):
# ------------------------------------

This paper assesses the impact of homeostasis on simulated model neurons receiving inputs that are either oscillatory, transiently bursting, or tonic. The authors find that increasing excitatory oscillatory input results in a homeostatic response that, when reaching an appropriate level of mean calcium, overcompensates the postsynaptic neuron's spiking response. That is, paradoxically, an increase in drive when combined with homeostasis results in a decreased response.

I think the main effect-which according to the authors does not occur with inhibitory oscillatory drive-is of interest and worth exploring more thoroughly. Under what conditions does it occur? What types of calcium channels are essential, what are inessential? Does spiking threshold or leak conductance matter? As it stands though, the paper provides one example of the effect without enough corresponding examples to allow the reader to understand precisely how the effect arises and why. I think I know the reason, but even a statement such as the one in the opening paragraph of this review is not present in the text. I suspect some simple controls, perhaps comparing high-threshold calcium channels versus low-threshold calcium channels as the main conveyors of calcium entry in the soma, would help provide some insight into the nature of the observed phenomenon.

> The reviewer's summary of our work was quite apt and covered the key mechanism nicely. (Better than we had, in truth). We added a similar summary to the opening of the Results section. It now reads,

> "We assessed the impact of homeostasis in a simple model that received oscillatory inputs that are either transiently bursting, or tonic. We found that increasing the strength of AMPAergic oscillatory inputs resulted in a homeostatic response that, when reaching an appropriate level of average calcium, overcompensates to actually suppresses the postsynaptic neuron's spiking. That is, paradoxically, an increase in drive when combined with homeostasis results in a decreased response."

Some of the ways the authors discuss the effect are a little counterintuitive - as though the behavior of the network is in place so as to become compatible with homeostasis, rather than the generally understood notion that homeostasis is present to ensure the network is stable. If it really is the general case that excitatory-input oscillations are incompatible with homeostasis then it is fine to generalize in this way, but a single example is insufficient-I can imagine other ways in which oscillations arising from excitatory input remain stable in the presence of homeostasis (I imagine it is all channel-type-dependent and parameter-dependent). It may be worth demonstrating that the homeostatic rhythmic model of O'Leary and Marder also fails with rhythmic excitatory input.

> We added a paragraph to the discussion to try and address the role of homeostasis, and its interaction with oscillations. It tries to make clear a key point: homeostasis and oscillations might often be destabilizing but this is because oscillations **share** synapses/dendrites with stimulus drives. Other modulators, ACh, DA, etc, use separate synapses and so have some degrees of independent freedom.

> "Homeostasis is generally viewed as a stabilizing force. This is the goal of homeostasis in our model as well. However oscillatory modulation is a unique form of modulation; it shares neurotransmitters, and even synapses, with non-modulatory stimuli. As a result, intrinsic homeostatic corrections for modulation must affect both the modulator and the driver \cite{Sherman1998}. Our work suggests this trade-off can be quite strong. To avoid this trade-off, sometimes anyway, we speculate the long-time scale of evolution may favor different types of oscillation for different functional roles. "

> To argue again for generality, we also added this explanation to beginning of the results section. It reads,

> "The homeostatic response in our simple model decreases the Sodium conductance and increases the Potassium conductance (Figure~\ref{fig:f1}\textbf{c}, \textit{bottom panel}). Combined these make it harder for the model cell to generate action potentials. This basic mechanism explains the loss of excitability in some conditions of our model and, in turn, the desynchronization of population activity."

> And to make the limitations of our, or any other, simple model clear we added a much stronger set of caveats to the existing limitations section in the discussion. It reads,

> "The nature of our model--that fact we use point neurons with only 6 currents--or the fact that our model is strictly feed-forward--without lateral or recurrent connections--means we don't know with confidence to what degree our model's effects will appear in more complex models, or in real neural systems. O'Leary \cite{OLeary2014} and others have shown how the homeostatic response can be exceptionally complex, and even unpredictable. It is therefore likely some neurons will have a homeostatic response that is nothing like that in our model. There may be many such examples."

> "We do know, however, that oscillations are a ubiquitous feature of cortical activity, as is Ca\textsuperscript{2+}-mediated intrinsic homeostasis. This means the ingredients for oscillation and homeostasis to interact are omnipresent in both sub-cortical and cortical areas. This response might be again very different than what our models show, but that there could be a broad range of possible effects, making our initial results in a simple model all the more intriguing."

> We also added the below text to our results about the artificiality of AMPA-theta in hippocampus, and more on its limits and generality.

> Of course theta oscillations in hippocampus are GABAergic, not AMPAergic as we model here. Exploring this artificial oscillation lets us show the central principle we want to highlight in our simple model. It seems that tonic oscillations--when joined with homeostasis--can profoundly alter the coding properties of what one would expect from oscillatory modulation. The degree of effect depends on the degree to which the oscillation changes the cell's tonic Ca$^{2+}$. It also shows that because of homeostasis, bursts are \textit{qualitatively distinct} from their tonic counterparts.

> The specific effect we observed (suppression of excitability) is seen because the homeostatic equations that \textit{we chose} respond to changes in Ca\textsuperscript{2+} by decreasing the conductance of the Na and Ca channels, and increasing the conductance of K and KCa channels (Eq~\ref{eq:dgdt}). The net effect of these dynamics is to decrease the sodium channel conductance, and increase the potassium. This makes firing an action potential less likely, for any given input. These channel dynamics are visualized in Figure~\ref{fig:f1}\textbf{c} and the corresponding FI curve is shown in Figure~\ref{fig:f4}.

> The we general effect we would expect for some other selection of channels, chosen by some other theorists or by evolution itself, depends on those channels. O'Leary \cite{OLeary2014} has shown how complex the relationship is between homeostasis, channel conductance, and neuronal firing patterns. It's complex enough that general statements are probably impossible. One one hand though our simple model has been shown by others to recapitulate hippocampal pyramidal cell's activity in laboratory settings \cite{LeMasson1993}. On the other, it is \textit{very simple} and so most probably incomplete at best. Experimental testing is needed.

A final general point - I find the writing rather unpolished in places, with non-grammatical sentences, and recommend careful editing in any future version.

> Noted.

Specific comments:

l.48 "kinetics of ion channels": I have not seen evidence of channel kinetics changing, only expression levels, so please cite the appropriate evidence.

l.58 An increase in threshold is not paradoxical, but, rather, the standard, expected homeostatic response.

> Correct. We updated the draft to reflect this.

l.82 On the "radical difference between real and simulated timescales": there is no clear justification for simulating a timescale that is radically different from the real one. In the absence of such justification, I recommend using the correct timescale.

> Unfortunately, computationally it is impractical to use the real one. The simulations would take many months to complete. We have added a more complete justification for this (necessary) choice. It includes a quantification of the differences.

> "In real systems, intrinsic homeostasis is thought to happen over minutes or days. However, simulation times that are hours or days long are not computationally feasible. So we follow the field and study a model where Ca\textsuperscript{2+} dynamics happen with a 4-second half-life, denoted by $\tau_h$. This might seem like a huge difference. But all that matters mathematically is that Ca\textsuperscript{2+} dynamics happen much slower than all the other synaptic/membrane dynamics. A timescale of $\tau_h > 4$ seconds is a reasonable choice to ensure this. The other major dynamics in the model are synapses with $< 30 ms$ half-lives. This leads to a 133 fold difference \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,Marder2014,OLeary2014,LeMasson1993,Abbott1993}."

l. 232 the background rate, r_b is not used in any equations so hard to tell which conditions and under what circumstances it is ever used. Also I don't see r_s used in the equations in the methods.

> The background rate was removed from the model prior to publication. Its inclusion in the paper was a mistake.

Minor points:

l. 27 "this quench" (invert) 
l.114 "change in Ca..." 
l.118 "effect" written twice 
l.129 "increase or decrease" 
l. 148 comma after oscillations. 
l. 167 "that" is wrong 
l. 170 "For example .." this is not a complete sentence.

> Thank you.

l.227 you have not specified which cells in the model comprise the hippocampal population

> Thank you. This is was ambiguous several places in the main text, and we have corrected it throughout.

l.230 The math equation looks like you are dividing by the parentheses instead of multiplying. I would put parentheses around r_0/2 to be clear.

> Thank you.

Reviewer #3 (Comments to the Author):
# ------------------------------------

In this paper, the authors add calcium-dependent homeostatic mechanisms to a feedforward network and find a paradoxical effect for tonic oscillations that are synaptically excitatory. This seems interesting, but after some effort, I found that I was unable to parse and evaluate and kept trying to ‘fill in the blanks’. In essence, my main comments are that I am unclear about various details (sorry if I missed them somewhere). For example:

> To clarify the larger questions the reviewer raised about the methods were moved and centralized in a short description of the model at the start of the results. It answers these questions:

1) what is the connectivity (if any) of the 100 HH neurons? (and how does it affect results?) 2) how does the connectivity and synaptic weights (from N to M as given in methods) affect results?
4) confusing/incomplete descriptions – e.g., they say that the ‘stimulus population is modeled by a fixed rate of 6 Hz’, but is this the rate that is changing in the figures? I am not sure what (parameters) are being changed (and how) when they say (line 77) the oscillatory strength, duration... The methods refer to a uniform distribution.
5) line 95/96: they say that “...increasing the strength of the excitatory oscillation increases excitability leading to an increase in the population firing rate”. I am confused – I thought they said that they are using ‘population firing rate’ as a measure of ‘excitability’ (bottom of p.2). Also, as per comments above, I am not sure exactly what is being done when they say that they are ‘increasing the strength of the excitatory oscillation’

> It reads,
> "We study Ca\textsuperscript{2+}-mediated homeostasis in a feed-forward population of 100 Hodgkin-Huxley neurons. By feedforward we mean neurons in the population are driven only by the inputs to the model. Cells have no lateral or recurrent connections. We modulate this population with neural oscillations. Oscillations are modelled as a set of 1000 Poisson neurons. To change the level of modulation strength, we changed the firing rate of the Poisson population. Average synaptic strengths were fixed, but drawn from a uniform distribution for each simulation (see \textit{Methods}). An oscillation's temporal duration was simulated as either tonic, which lasted for the entire experiment, or as a 4-cycle burst. Bursts were delivered only at the end of a trial, where they overlap with the input stimulus (Figure~\ref{fig:f1}\textbf{a}). The stimulus input was also drawn from a 1000 cell Poisson population. It lasted 0.5 seconds, and had an average firing rate of 6 Hz. The stimulus was fixed for all experiments, and was delivered only once homeostatic equilibrium is reached \cite{Barth2012} (Figure~\ref{fig:f1}). During the stimulus window we measure the population's firing rate and synchrony. It is these number we report throughout."

3) incomplete/inconsistent parameters – e.g., what inhibitory reversal potential used? Is V_e in aq1 `*-Table same as E_s in eqn 3?

> We corrected the inconsistencies between the parameter table and equations in the methods. How embarrassing!

6) what is meant by the ‘firing threshold’ (line 59). That is, please define/specify what is intended to avoid confusion/circularity.

> The firing threshold is, it turns out, not well defined in the literature. Especially for modelling and theory. Sidestepping this, we now refer just to excitability in the paper. We can measure excitability directly using classic FI-curves, and link it directly to biophysical changes via Na/K conductances.

Once this is clarified, it seems to me that the authors should provide some insight to their paradoxical result (or at least explanation by doing enough numerical explorations) E.g., are the differences a matter of different driving forces with excitatory and inhibitory?

> FI-curves also provide clear a way to demonstrate the driving force for AMPAergic suppression of firing. We show this in the new Figure 4, wherein small square waves of current are injected into each cell's ``soma''. The number of action potentials which follow are collected and averaged. The more excitable the cells, the higher this Pop. rate will be.

Minor: line 88 – not a sentence – “Both of which...”

> Thank you.

### 2020-03-06 main.tex
Reviewer response 1 (JNP) complete

@@
 \begin{abstract}
-Neural oscillations are observed ubiquitously in the mammalian brain.
-However the stability of oscillations is highly variable.
-Some oscillations are \textit{tonic}, lasting for seconds or even minutes; others are unstable, appearing only as a single-cycle \textit{burst}.
-In a model of hippocampal neurons, we use numerical simulations to show how these different forms of rhythm stability can interact with activity-dependent homeostasis to profoundly alter the modulatory effect of neural oscillations.
-Under homeostasis, tonic oscillations that are synaptically excitatory have a paradoxical effect; they decrease excitability and desynchronizing firing.
-Tonic oscillations that are synaptically inhibitory--like those in a real hippocampus--fail to generate new action potentials and so provoke no homeostatic response.
-This may explain why the theta rhythm in hippocampus is synaptically inhibitory: inhibitory oscillations don't raise the firing threshold, as excitatory oscillations do, and so can preserve each cell's dynamic range.
-Based on these simulations, we also speculate that homeostasis may explain why excitatory intra-cortical and intra-layer oscillations often appear as bursts.
-In our model bursts minimally interact with the slow homeostasis time constant and so retain typical excitatory effects.\\
+Neural oscillations are observed ubiquitously in the mammalian brain, but their stability is quite variable.
+Some oscillations are \textit{tonic} and last for seconds or even minutes.
+Others oscillations are unstable.
+They appear for only as a single-cycle \textit{burst}.
+Likewise, some oscillations rely on excitatory AMPAergic synapses, but others are GABAergic and inhibitory.
+Why both kinds of diversity are present is not known.
+We hypothesized Ca$^{2+}$-dependent homeostasis is important to finding an explanation.
+We tested this in a highly simplified model of hippocampal neurons.
+In this model homeostasis can profoundly alter the modulatory effect of neural oscillations.
+Under homeostasis, tonic AMPAergic oscillations actually \textit{decrease} excitability and desynchronize firing.
+Tonic oscillations that are synaptically GABAergic--like those in real hippocampus--don't provoke a homeostatic response, however.
+If our simple model is correct, homeostasis can explain why the theta rhythm in hippocampus is synaptically inhibitory: GABA has little to no intrinsic homeostatic response, and so can preserve a cell's dynamic range under tonic oscillation.
+Based on these results we also speculate that homeostasis can explain why AMPAergic oscillations in cortex often appear as bursts.
+Bursts minimally interact with the slow homeostasis time constant, and so retain their excitatory effect.\\
 \end{abstract}
@@
 \noindent
-\textsc{\textbf{New and Noteworthy}} In a model of hippocampal neurons we report a paradoxical result: Ca\textsuperscript{2+}-mediated homeostasis causes normally synchronizing AMPAergic oscillations to become inhibitory and desynchronizing.
-We conjecture this could be why the theta rhythm is inhibitory.
-In pyramidal cells GABA doesn't generate new action potentials and so provokes no homeostatic response.
-The intricate interplay of neuromodulators, like acetylcholine, with homeostasis is well known.
-The interplay between oscillations and homeostasis might be as complex, and as important.
+\textsc{\textbf{New and Noteworthy}} The intricate interplay of neuromodulators, like acetylcholine, with homeostasis is well known.
+The interplay between oscillatory modulation and homeostasis is not.
+We studied oscillatory modulation and intrinsic homeostasis for the first time in a simlified model of hippocampal neurons.
+We report a paradoxical result: Ca\textsuperscript{2+}-mediated homeostasis causes normally synchronizing AMPAergic oscillations to become inhibitory and desynchronizing.
+This result, along with other new observations, suggests homeostasis might be just as complex -- and just as important -- for oscillations as it is for all other kinds of modulation.
 
@@
 Modulation perturbs excitability, which we define as the propensity for a stimulus to elicit an action potential.
-Homeostasis acts to this quench perturbation, driving the excitability of the cell back to a biologically desirable set point \cite{LeMasson1993,Abbott1993}.
-While the interplay between chemical modulators and homeostasis has been studied for more than 20 years \cite{LeMasson1993,Abbott1993,Golowasch1999,Marder2014,Marder2015,Gutierrez2013}, the relationship between network-level synaptic modulations--like neural oscillations--and homeostasis is not well understood, on either theoretical or empirical grounds.
+Homeostasis acts to quench this perturbation, driving the excitability of the cell back to a biologically desirable set point \cite{LeMasson1993,Abbott1993}.
+While the interplay between chemical modulators and homeostasis has been studied for more than 20 years \cite{LeMasson1993,Abbott1993,Golowasch1999,Marder2014,Marder2015,Gutierrez2013}, the relationship between population-level synaptic modulations--like neural oscillations--and homeostasis is not well understood, on either theoretical or empirical grounds.
 Like neuromodulators, neural oscillations alter excitability and firing statistics.
-Uniquely, though, oscillations create synchronous windows of activity \cite{Lisman2013,Voytek2015}.
-Temporally grouping action potentials into windows improves signal to noise and increases the number of coincident firing events \cite{Chen2013,Zhou2015,Voytek2015a,Peterson2017}, driving learning at individual synapses \cite{Muller2011,Song2000,Markram1997}.
+Uniquely though, oscillations can group action potentials into synchronous windows of activity \cite{Lisman2013,Voytek2015}.
+Grouping action potentials can improve signal to noise and increase the number of coincident firing events \cite{Chen2013,Zhou2015,Voytek2015a,Peterson2017}, which drives learning at individual synapses \cite{Muller2011,Song2000,Markram1997}.
 
-We propose the same mechanisms that link homeostasis with chemical neuromodulation can also come into play during oscillatory modulation.
-After all, both kinds of modulation lead to tonic changes in spiking and in Ca\textsuperscript{2+} concentration, which can in turn drives changes to intrinsic homeostasis \cite{Liu1998}.
+We propose the same mechanisms that link homeostasis with chemical neuromodulation can also come into play with oscillatory modulation.
+After all, both kinds of modulation lead to tonic changes in spiking and, as a result, changes in Ca\textsuperscript{2+} concentration \cite{Liu1998}.
 
 To test this, we model activity-dependent intrinsic homeostasis in a feed-forward population of hippocampal pyramidal cells \cite{Siegel1994}.
-In this model homeostasis is mediated by a Ca\textsuperscript{2+}-dependent mechanism \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,OLeary2014} pioneered by LeMasson \cite{LeMasson1993,Abbott1993}.
-In this model, Ca\textsuperscript{2+} acts a sensor or proxy for tonic changes in the membrane voltage.
+By feed-forward we mean there are no lateral or recurrent connections in the population.
+Firing activity is driven only by afferent synaptic input, and homeostasis is mediated by a Ca\textsuperscript{2+}-dependent mechanism \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,OLeary2014} pioneered by LeMasson \cite{LeMasson1993,Abbott1993}.
+Here Ca\textsuperscript{2+} acts a sensor for tonic changes in the membrane voltage.
 To counter tonic changes in Ca\textsuperscript{2+} levels, the expression of ion channels is altered, returning the Ca\textsuperscript{2+} level to a predefined ``good'' value \cite{Golowasch1999,OLeary2013}.
@@
 To minimize the effect of homeostasis on the rise and fall of action potential dynamics, we also add a KCa channel which is not present in Siegel \citep{Siegel1994}.
-In real cells intrinsic homeostasis changes the expression level, membrane trafficking, and kinetics of ion channels.
+In real cells intrinsic homeostasis changes the expression level of ion channels \cite{OLeary2013}.
 Here, these details are not directly simulated.
-Instead we mimic the net or bulk effect of all these changes by altering the maximum conductance of individual ionic channels, which are modelled using Hodgkin-Huxley terms \cite{LeMasson1993,OLeary2013,OLeary2014}.
+Instead we mimic the net or bulk effect of all ion channels using a single equation \cite{LeMasson1993,OLeary2013,OLeary2014}.
 
-Given that chemical modulators operate on both short and long times scales \cite{Marinelli2014,Marder2014,Cohen2015,Daw2002}, we examined two timescales of oscillation.
-We contrast the effects of shorts bursts of oscillation to long lasting tonic rhythms, both of which are observed {\textit{in vivo}}, and which may play different physiological, cognitive, and computational roles \cite{Lundqvist2016,vanEde2018}.
-We also explore how oscillation duration interacts with synapse type, examining both AMPA- or GABA-ergic oscillations.
+Given chemical modulators operate on both short and long times scales \cite{Marinelli2014,Marder2014,Cohen2015,Daw2002}, we examined two timescales of oscillatory modulation.
+We contrast the effects of short 4-cycle bursts of oscillation to long lasting tonic rhythms.
+Both are observed {\textit{in vivo}}, and are known to play different physiological, cognitive, and computational roles \cite{Lundqvist2016,vanEde2018,Peterson2017}.
+We also explore synapse type, examining both AMPA- or GABA-ergic oscillations.
 
-In our highly simplified model of hippocampal cells, intrinsic homeostasis can profoundly alter the modulatory effect of neural oscillations.
-\textit{Tonic} excitatory modulation, paradoxically, generates a homeostatic response that increases the firing threshold.
-This suppresses excitability which, in turn, desynchronizes population activity.
-\textit{Bursts} of excitation, meanwhile, don't show homeostatic suppression, and may even benefit from a weak level of homeostasis.
-This is due the constant noisy background level input also present in the model, which homeostatic mechanisms help control.
-Inhibitory GABAergic oscillations, however, show little to no homeostatic effect, suggesting that inhibitory oscillations might better isolate any phase coding scheme from the stimulus-driven response.
+% \textit{Bursts} of excitation don't show homeostatic suppression.
+They are two fast for the slow Calcium sensor to detect.
+Inhibitory GABAergic also have little to no homeostatic effect, suggesting that inhibitory oscillations might better isolate any phase coding scheme from the stimulus-driven response.
 This is might be important, for example, in hippocampal phase-coding schemes of memory \cite{Lisman2013}.
@@
 \section*{Results}
-We study Ca\textsuperscript{2+}-mediated homeostasis in a feed-forward network, using single compartment neurons.
-We modulate this network using neural oscillations.
-Oscillations are simulated either as tonic, lasting for the entire experiment, or may as a burst at the end, where it overlaps with input stimulus (Figure~\ref{fig:f1}\textbf{a}).
-Input into the model has both a constant background level of noise (not shown), and a stronger stimulus that is delivered once homeostatic equilibrium is reached \cite{Barth2012} (Figure~\ref{fig:f1}).
-During this strong stimulus we measure the network's response and how population firing and synchrony are affected by homeostasis under different oscillatory regimes.
+We assessed the impact of homeostasis in a simple model that received oscillatory inputs that are either transiently bursting, or tonic.
+We found that increasing the strength of AMPAergic oscillatory inputs resulted in a homeostatic response that, when reaching an appropriate level of average calcium, overcompensates to actually suppresses the postsynaptic neuron's spiking.
+That is, paradoxically, an increase in drive when combined with homeostasis results in a decreased response.
 
-Each experiment began with an instantiation of a randomly generated feed-forward network.
-A single neuron from this network is depicted in Figure~\ref{fig:f1}\textbf{b}.
-This network is subjected to a range of modulatory and control conditions, including oscillatory strength, duration, and synapse type (AMPA or GABA).
+Specifically, we study Ca\textsuperscript{2+}-mediated homeostasis in a feed-forward population of 100 Hodgkin-Huxley neurons.
+By feedforward we mean neurons in the population are driven only by the inputs to the model.
+Cells have no lateral or recurrent connections.
+We modulate this population with neural oscillations.
+Oscillations are modelled as a set of 1000 Poisson neurons.
+To change the level of oscillatory modulation strength, we alter the firing rate of the Poisson population.
+Average synaptic strengths were fixed, but drawn from a uniform distribution for each simulation (see \textit{Methods}).
+An oscillation's temporal duration was simulated as either tonic, which lasted for the entire experiment, or as a 4-cycle burst.
+Bursts were delivered only at the end of a trial, where they overlap with the input stimulus (Figure~\ref{fig:f1}\textbf{a}).
+The stimulus input was also drawn from a 1000 cell Poisson population.
+It lasted 0.5 seconds, and had an average firing rate of 6 Hz.
+The stimulus was fixed for all experiments, and was delivered only once homeostatic equilibrium was reached \cite{Barth2012} (Figure~\ref{fig:f1}).
+During the stimulus window we measured the population's firing rate and synchrony.
+It is these numbers we report throughout.
+
+Each experiment began with an instantiation of a randomly generated population.
+A single example neuron is depicted in Figure~\ref{fig:f1}\textbf{b}.
+This population is subjected to a range of modulatory and control conditions, including oscillatory strength, duration, and synapse type (AMPA or GABA).
 Each experiment lasted 20 seconds.
 In Figure~\ref{fig:f1}\textbf{c} we depict key aspects of model output during an experiment.
-In real systems, intrinsic homeostasis is thought to happen over minutes or days.
-However, simulation times that are hours or days long are not computationally feasible.
-As a result, we follow the field and study a model where Ca\textsuperscript{2+} dynamics happen with a 4-second half-life, denoted by $\tau_h$.
-Despite the radical difference between real and simulated time-scales, all that matters mathematically is that Ca\textsuperscript{2+} dynamics happen much slower than all the other synaptic/membrane dynamics.
-In practice, this means a timescale of $\tau_h > 4$ seconds is a reasonable first-order approximation \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,Marder2014,OLeary2014,LeMasson1993,Abbott1993}.
 
 After homeostatic equilibrium is reached, we measure two features: the synchrony between action potentials (measured by the Kappa correlation) and changes in the excitability of the system (measured as a change in population firing rate).
-Both of which are defined in the \textit{Methods}.
+Both of these measures are defined in the \textit{Methods}.
 To ensure a consistent comparison between experiments, measurements were made over the same 4-cycle or 0.5 second period in all simulations.
@@
 
-% At the end of each experiment we measure three main features: synchrony between action potentials between different neurons, for a given type of modulation; changes in the excitability of the system--measured as a change in population firing rate--and; the ``computational error'' introduced by the combination of modulation and homeostasis.
-Error is defined as the mean squared error (MSE) measured by comparing spike-times recorded during experimental conditions to recordings made in identical populations simulated without modulation/homeostasis.
-To ensure a consistent comparison between experiments, measurements were made over the same 4-cycle period \emph{for all simulations}.
+The homeostatic response in our simple model decreases the sodium conductance and increases the potassium conductance (Figure~\ref{fig:f1}\textbf{c}, \textit{bottom panel}).
+Combined these make it harder for the model cell to generate action potentials.
+This basic mechanism explains the loss of excitability in some conditions of our model and, in turn, the desynchronization of population activity.
 
@@
 Top panel depicts the stimulus, applied over the last 0.5 seconds of each simulated trial.
-Middle and bottom panels depict the two modes of oscillation we examined--tonic and bursting.
+Middle and bottom panels depict the two modes of oscillation we examined: tonic and bursting.
     \textbf{b}. Illustration of a single model neuron, its major currents, its inputs (top arrow) and its output (bottom arrow).
@@
 
+Note that in real systems intrinsic homeostasis is thought to happen over minutes or days.
+However, simulation times that are hours or days long are not computationally feasible.
+So we follow the field and study a model where Ca\textsuperscript{2+} dynamics happen with a 4-second half-life, denoted by $\tau_h$.
+This might seem like a huge difference.
+But all that matters mathematically is that Ca\textsuperscript{2+} dynamics happen much slower than all the other synaptic/membrane dynamics.
+A timescale of $\tau_h > 4$ seconds is a reasonable choice to ensure this.
+The other major dynamics in the model are synapses with $< 30 ms$ half-lives.
+This leads to a 133 fold difference \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,Marder2014,OLeary2014,LeMasson1993,Abbott1993}.
+
 % -----------------------------------------------------------------------
@@
 Homeostasis completely inverts the effect of tonic AMPAergic oscillatory modulation.
-To see how, first consider how the network responds to a tonic excitatory modulation without homeostasis.
-Without homeostasis, increasing the strength of the excitatory oscillation increases excitability, leading to an increase in the population firing rate.
-Additionally, the neuromodulatory effect of the oscillations is such that action potentials are grouped by oscillation phase, resulting in an increase in population synchrony (Figure \ref{fig:f2}\textbf{b}, black line).
+These excitatory oscillations now result in decreases in both population firing rate, and population synchrony.
+To see how, first consider how the population responds to a tonic excitatory modulation \textit{without} homeostasis.
+Without homeostasis, increasing the strength of the excitatory oscillation increases excitability and increases the population firing rate.
+The action potentials are also grouped by oscillation's rising phase, increasing the population's synchrony (Figure \ref{fig:f2}\textbf{b}, black line).
+
 With homeostasis this pattern inverts.
@@
 The stronger the oscillation, the more suppressive the result.
-By the time the oscillation is about half the strength of the stimulus (which we fix at a firing rate of 6 Hz) the stimulus is completely suppressed (\textit{i.e.}, the population firing rate approaches 0) (Figure \ref{fig:f2}\textbf{c}).
+By the time the oscillation is about half the strength of the stimulus (which we fix at a firing rate of 6 Hz) the stimulus is completely suppressed and the population firing rate approaches 0 (Figure \ref{fig:f2}\textbf{c}).
 
 We compared the effect of tonic oscillation to short 4-cycle bursts of excitatory modulation, presented only during the stimulus.
-Here, the oscillation period is far too short to engage any additional homeostasis.
+Here, the oscillation period is far too short to provoke homeostasis.
 This means that increases in oscillatory strength continue to increase population firing rate and synchrony (Figure~\ref{fig:f2}\textbf{e} and \textbf{f}).
 
-Our model suggests tonic oscillations can profoundly alter coding properties of synaptically excitatory oscillations.
-This means that bursts of excitatory oscillation are \textit{qualitatively distinct} from their tonic counterparts.
-The decrease in excitability caused by tonic oscillation can be explained by a direct homeostatic change.
-Larger tonic oscillations lead to larger tonic increases in the membrane potential, which in turn raise Ca\textsuperscript{2+} levels.
-The homeostatic equations respond to this change Ca\textsuperscript{2+} by decreasing the conductance of the Na and Ca channels, and increasing the conductance of K and KCa channels (Eq~\ref{eq:dgdt}).
-The net effect of these dynamics is a increase in the firing threshold.
+Of course theta oscillations in hippocampus are GABAergic, not AMPAergic as we model here.
+Exploring this artificial oscillation lets us show the central principle we want to highlight in our simple model.
+It seems that tonic oscillations--when joined with homeostasis--can profoundly alter the coding properties of what one would expect from oscillatory modulation.
+The degree of effect depends on the degree to which the oscillation changes the cell's tonic Ca$^{2+}$.
+It also shows that because of homeostasis, bursts are \textit{qualitatively distinct} from their tonic counterparts.
 
-The inversion in Kappa seen in Figure~\ref{fig:f2}\textbf{d} as well as in \ref{fig:f4}\textbf{b}, is an artifact of the decrease in excitability; it's a low effect $N$ effect.
-As population spiking becomes more strongly suppressed the total number of spikes declines to the point where the bins used to calculate Kappa often contain no spikes.
-This in turn inflates Kappa values.
+The specific effect we observed (suppression of excitability) is seen because the homeostatic equations that \textit{we chose} respond to changes in Ca\textsuperscript{2+} by decreasing the conductance of the Na and Ca channels, and increasing the conductance of K and KCa channels (Eq~\ref{eq:dgdt}).
+The net effect of these dynamics is to decrease the sodium channel conductance, and increase the potassium.
+This makes firing an action potential less likely, for any given input.
+These channel dynamics are visualized in Figure~\ref{fig:f1}\textbf{c} and the corresponding FI curve is shown in Figure~\ref{fig:f4}.
+
+The general effect we would expect for some other selection of channels, chosen by some other theorists or by evolution itself, depends on those channels.
+O'Leary \cite{OLeary2014} has shown how complex the relationship is between homeostasis, channel conductance, and neuronal firing patterns.
+It's complex enough that general statements are probably impossible.
+On the one hand, our simple model has been shown by others to recapitulate hippocampal pyramidal cell's activity in laboratory settings \cite{LeMasson1993}.
+On the other, it is \textit{very simple} and so most probably incomplete at best.
+Experimental testing is needed.
+
+Note that the inversion in Kappa seen in Figure~\ref{fig:f2}\textbf{d} as well as in \ref{fig:f5}\textbf{b}, is the result of the decrease in excitability.
+Put another way, it's a low $N$ effect.
+As population spiking is suppressed, the total number of spikes declines to the point where the bins used to calculate the Kappa correlation often contain no spikes.
+This leads to a somewhat misleading increase in synchrony.
 
 \subsection*{Inhibitory modulation.}
-Both sustained tonic and bursting inhibitory oscillatory modulation do not lead to a homeostatic response in our model.
+GABAergic oscillatory modulation does not lead to a homeostatic response in our model.
+This is true for both tonic oscillations and bursts.
 In all cases, as oscillatory strength increases, population firing declines dramatically (grey lines in Figure \ref{fig:f2}\textbf{a-d} and light yellow lines in panel \textbf{e-f}).
-
-% \subsection*{The synaptic nature of hippocampal theta.}
-
-% \subsection*{Modulation, homeostasis, and error.}
-% When increasing or decreasing the firing rate of a population, or its synchrony, action potentials can be added or re-arranged arbitrarily.
-We've suggested previously that an efficient modulator is one that adds or moves spikes as little as possible, while achieving a set level of effect \cite{Peterson2018a}.
-We define spike-time deviations introduced by modulation as an error compared to the original, unmodulated, population output (see Eq~\ref{eq:error} and \cite{Peterson2018a} for more background).
-We compare observed errors with and without homeostasis, for burst and tonic excitatory modulation.
-Given that synaptic inhibitory oscillatory modulation failed to trigger any homeostatic effects we don't include it in these analyses.
-
-% Under homeostasis, tonic excitatory oscillations suppress the network responses to stimulation, leading to extremely high errors compared to an identical model simulated without homeostasis included (Figure~\ref{fig:f3}\textbf{a}).
-Bursts of excitatory oscillatory modulation, however, actually \textit{reduce} error compared to their non-homeostasis counterpart (Figure~\ref{fig:f3}\textbf{a}.).
-This decrease in error is driven by the baseline level of noise present in our stimulus model, which when input into the hippocampal population leads to a slight reduction in the excitability, this then positively interacts with the modulatory burst which ends more effective modulator compared to the baseline--homeostasis free--model.
 
@@
 
-% \begin{figure}
-% \centering
-% \includegraphics[width=0.6\textwidth]{fig3.png}
-% \caption{\label{fig:f3}
-%     Synchrony versus error during excitatory modulation, with (black) and without (grey) homeostasis.
-%     \textbf{a.} Tonic modulation.
-%     \textbf{b.} 4-cycle burst.
-% }
-% \end{figure}
-
-% -----------------------------------------------------------------------
-\subsection*{The effect of Ca\textsuperscript{2+} concentration.}
-When the model is run with only stimulus-driven homeostasis, the Ca\textsuperscript{2+} concentration equilibrates to about 0.003 mM.
-We used this as a standard target value for all modulation experiments, until now.
-When we vary this value in 0.0002 mM increments, population rate and synchrony either increases or decreases depending on whether the Ca\textsuperscript{2+} increases or decreases, shown Figure \ref{fig:f4}\textbf{a-b}.
-However despite different initial Ca\textsuperscript{2+} concentrations, each model still shows an identical set of trends as the strength of the oscillation increases (Figure \ref{fig:f4}).
-That is, increasing or decreasing the target concentration shifts the overall excitability of the network, in an approximately linear way.
-This means that while the initial choice of 0.003 mM was arbitrary, the qualitative pattern of results we report is not dependent on this choice.
+In Figure~\ref{fig:f2} we compared AMPA and GABAergic models of the theta rhythm (8 Hz).
+This is done to illuminate their differences, but is not biologically realistic.
+Physiological theta is a GABAergic.
+AMPAergic oscillations of the hippocampus are, however, seen in the high gamma range (60-90 Hz).
+In Figure~\ref{fig:f3} we confirm AMPAergic oscillations in this range follow the same pattern as theta.
+For both 60 and 90 Hz gamma, tonic oscillations suppress excitability and desynchronize firing.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig4.png}
+\includegraphics[width=0.5\textwidth]{fig3.png}
+\caption{\label{fig:f3}
+    The effect of fast AMPAergic gamma oscillations for both tonic oscillation and 4-cycle bursts. 
+    Top panel (\textbf{a}.) is the observed population firing rate averaged over a 0.05 second or 0.033 second stimulus.
+Bottom (\textbf{b}.) is synchrony over the same period.
+Columns in both \textit{a.} and \textit{b.} represent with 60 or 90 Hz gamma.
+The stimulus window was matched to the length of time 3 cycles of oscillation would take.
+}
+\end{figure}
+
+\subsection*{Homeostasis and excitability.}
+In our model Ca\textsuperscript{2+} influx is driven by increases in the membrane voltage, and Ca\textsuperscript{2+} concentration is restored both by a passive and linear recovery (Eq~\ref{eq:passive}) and active homeostasis (Eq.~\ref{eq:dgdt}).
+Homeostasis works by negative feedback.
+Increases in Ca\textsuperscript{2+} concentration decrease the conductance of both the sodium and potassium channels in our model (Eq.~\ref{eq:HHH}).
+This decrease in conductance in turn lowers the membrane voltage down closer to its resting value.
+As the voltage falls so does the Ca\textsuperscript{2+} concentration.
+The excitability changes we see with homeostasis must be due to the decrease in Na conductance.
+Changes due to K conductance--which acts to recover the voltage following a spike--can't alter baseline.
+
+One the simplest, classic ways to measure the excitability of a cell--simulated or real--is with an FI-curve.
+In an FI-curve a small square wave of current is injected into the cell, and the resulting increase in firing rate (if any) is averaged over a short window.
+In Figure~\ref{fig:f4} we show the average FI-curve for 100 simulated hippocampal neurons.
+Under homeostasis the FI-curve is lower, which means there has been a substantial suppression in the membrane excitability.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{fig4.png}
 \caption{\label{fig:f4}
-    Synchrony and excitability shift linearly with the target Calcium concentration, [Ca].
-Plotted here is the effect of Ca\textsuperscript{2+} on a tonic excitatory modulation.
-The black value (0.003 mM) is the standard reference value used in all previous simulations.
+Excitability and homeostasis.
+An average FI-curve for 100 hippocampal neurons, with (black) and without (grey) homeostasis.
+In both models, oscillation was tonic and AMPAergic.
+The square-wave pulse lasted 0.1 seconds, and ranged from 0.1 to 1 $\mu$amps, sampled every 0.01 $\mu$amps.
+}
+\end{figure}
+
+% -----------------------------------------------------------------------
+\subsection*{Controls and perturbations.}
+To ensure our parameter choices in the model were generalizable, we perturbed several key values by 10-20 percent.
+
+When the model is run with only stimulus-driven homeostasis, the Ca\textsuperscript{2+} concentration equilibrates to about 0.003 mM.
+We used this as a standard target value for all modulation experiments, until now.
+When we vary this value in 0.0002 mM increments, population rate and synchrony either increases or decreases depending on whether the Ca\textsuperscript{2+} increases or decreases, shown Figure \ref{fig:f5}\textbf{a-b}.
+However despite different initial Ca\textsuperscript{2+} concentrations, each model still shows an identical set of trends as the strength of the oscillation increases (Figure \ref{fig:f5}).
+That is, increasing or decreasing the target concentration shifts the overall excitability of the population, in an approximately linear way.
+This means that while the initial choice of 0.003 mM was arbitrary, the qualitative pattern of results we report is not dependent on this choice.
+
+We chose a 4-cycle length for bursts somewhat arbitrarily.
+Real oscillations range in duration from single cycles to seconds of even minutes \cite{Lundqvist2016,vanEde2018}.
+Our burst/tonic distinction is therefore a false (but useful) dichotomy.
+To find a better estimate of how an oscillation would need to last, on average, to generate a homeostatic response we ran some long simulations using a more realistic homeostatic time constant ($\tau_h$ = 600 seconds; up from 20 seconds).
+It seems an oscillation would need to be present \textit{at least} 10\% of the time to generate a response (Figure~\ref{fig:f5}\textbf{c}).
+
+Real theta rhythms drift in frequency \cite{Buzsaki2015}.
+To test if small differences in frequency have a notable effect, we simulated small drifts in frequency of theta ranging from 6-10 Hz.
+Frequency drift had no visible effect on the homeostatic suppression of excitability (Figure~\ref{fig:f5}\textbf{d}-\textbf{e}).
+
+The leak conductance ($g_L$, Eq.~\ref{eq:HHH}) plays a key role in setting the resting potential of the cell.
+The resting potential in turn helps determines overall membrane excitability, and sets the resting Ca\textsuperscript{2+} concentration.
+This means our Ca\textsuperscript{2+} set point (of 0.003 mM) depends on the $g_L$ in practice but our choice of of 0.003 mM was arbitrary.
+To test if a small mis-match between $g_L$ and the set-point altered our results, we ran several experiments perturbing the leak conductance while keeping the Ca\textsuperscript{2+} set-point fixed.
+Changes to leak conductance had only a small effect on homeostasis (Figure~\ref{fig:f5}\textbf{f}-\textbf{g}).
+
+\begin{figure}
+\centering
+\includegraphics[width=1\textwidth]{fig5.png}
+\caption{\label{fig:f5}
+    Control experiments.
     \textbf{a.} Change in population hippocampal population rate for different levels of target [Ca] (colors) as a function oscillation strength (Osc.
@@
 However in choosing such as small value we follow the majority of the homeostasis modeling literature (for more on this see the \textit{Discussion}).
+    \textbf{d}.-\textbf{e}. The effect of frequency drift (see legend) on homeostasis with tonic AMPAergic modulation.
+    \textbf{f}.-\textbf{g}. The effect of altering the leak conductance ($g_L$) on homeostasis with tonic AMPAergic modulation.
 }
@@
 Our scientific understanding of homeostasis has been shaped as much by theoretical work as empirical \cite{Marder2014}.
-In an attempt to understand the interaction between oscillatory modulations and homeostasis, we begin by studying one of the simplest models used in early studies of homeostasis--a population of point neurons \cite{LeMasson1993}.
+In an attempt to understand the interaction between oscillatory modulations and homeostasis, we begin by studying one of the simplest models used in early studies of homeostasis--a population of point neurons in the hippocampus \cite{LeMasson1993}.
 
-The feed-forward model of hippocampal pyramidal cells studied here serves as a simple initial model to answer three basic questions.
-One, do tonic oscillations--generated with biologically-consistent parameters--engage homeostatic mechanisms?
+Our simple models offer answers to three basic questions.
+One, do tonic oscillations engage homeostatic mechanisms?
 Two, does homeostasis in turn change the oscillation's function?
 Three, do short bursts of oscillation have distinct effects from tonic oscillations?
-Put another way: can homeostasis help explain why some oscillations, such as hippocampal theta, tend to appear as tonic rhythms while other oscillations tend to appear as bursts?
-Our results suggest that homeostasis can explain why hippocampal theta is synaptically inhibitory, and why cortical oscillations often appear as bursts.
+Put another way: can homeostasis explain why some oscillations, such as hippocampal theta, tend to appear as tonic rhythms while other oscillations (like those in cortex) tend to appear as bursts?
+
+Homeostasis is generally viewed as a stabilizing force.
+This is the goal of homeostasis in our model as well.
+However oscillatory modulation is a unique form of modulation; It shares neurotransmitters, and even synapses, with non-modulatory stimuli.
+As a result, homeostatic corrections for modulation must affect both the modulator and the driver \cite{Sherman1998}.
+Our work suggests this trade-off can be quite strong.
+To avoid this trade-off, sometimes anyway, we speculate that the long-time scale of evolution may favor different types of oscillation for different functional roles.
 
@@
 
-While many neocortical oscillations tend to be bursty, a counterexample is visual cortical alpha, which can become tonic and high power when humans rest with their eyes closed.
-Even though both cortical and sub-cortical alpha generators are synaptically excitatory, this rhythm had been classically and paradoxically associated suppression of excitability \cite{Jensen2002,Bonnefond2012,Peterson2017}.
-While several competing explanations have been offered \cite{Bonnefond2012,Lange2013,Peterson2017} for this paradox, our work raises another possibility.
-Though we modelled hippocampal cells, the same Ca\textsuperscript{2+} homeostatic mechanisms exist in neocortex.
-Meaning strong and tonic alpha oscillations, combined with homeostasis, directly suppress population firing in visual cortex.
+An important counterexample to beta is visual cortical alpha (8-10 Hz).
+Alpha is tonic and high powered when humans rest with their eyes closed.
+Even though both cortical and sub-cortical alpha generators are synaptically excitatory, this rhythm has however been associated with suppression of excitability \cite{Jensen2002,Bonnefond2012,Peterson2017}.
+While several competing explanations have been offered \cite{Bonnefond2012,Lange2013,Peterson2017} for this, our work raises another possibility that is complementary to the others.
+Specifically, though we modelled hippocampal cells, the same Ca\textsuperscript{2+} homeostatic mechanisms exist in visual cortex (and the neocortex as a whole).
+This means that strong and tonic alpha oscillations, combined with homeostasis, should directly suppress population firing in visual cortex.
 Such an effect, were it to occur, would last well past the moment of oscillation offset.
-In fact, such long term effects of alpha have been reported in the literature, though the physiological mechanism was often unclear.
-Our work suggests that intrinsic homeostasis may underlie these effects.
+In fact, such long term effects of alpha have been reported in the literature \cite{Jensen2002,Bonnefond2012}, though the physiological mechanism was often unclear.
+Our work suggests that intrinsic homeostasis may underlie this effect.
 
 \subsection*{Limitations}
-The nature of our model--that fact we use point neurons with only 6 currents--or the fact that our model is strictly feed-forward--without lateral or recurrent connections--means we don't know with confidence to what degree our model's effects will appear in more complex models, or in real neural systems.
-For example, the primate or rat hippocampus.
+The nature of our model--the fact that we use point neurons with only 6 currents--or the fact that our model is strictly feed-forward--without lateral or recurrent connections--means we don't know with confidence to what degree our model's effects will appear in more complex models, or in real neural systems.
+O'Leary \citep{OLeary2014} and others have shown how exceptionally complex the homeostatic response is.
+It is therefore likely some neurons will have a homeostatic response that is nothing like that in our model.
+Others may be naturally immune to oscillatory effects.
+There may be many such examples.
+
 We do know, however, that oscillations are a ubiquitous feature of cortical activity, as is Ca\textsuperscript{2+}-mediated intrinsic homeostasis.
 This means the ingredients for oscillation and homeostasis to interact are omnipresent in both sub-cortical and cortical areas.
+This response might be very different than what we found.
+But that there could be broad range of effects makes our initial results in a simple model all the more intriguing.
 
-Homeostatic interactions depend on a number of factors specific to each cell and circuit.
+The basic mechanisms of Ca\textsuperscript{2+}-mediated intrinsic homeostasis are conserved across phyla \cite{Tran2017}.
+Meaning the qualitative properties of our model may also be present across phyla.
+That said the mechanisms for homeostatic interactions, along with their quantitative properties, depend on a number of factors specific to each cell and circuit.
 These include the duty cycle, power, and frequency of an oscillation, as well as on synaptic strengths and their location in the dendritic tree (an idea we return to below).
@@
 
-Synaptic homeostasis may also play a role in tuning a neuron's response to all types of oscillatory perturbations, as both excitatory and inhibitory synapses are susceptible \cite{Cannon2016}, though in the case of excitatory modulation synaptic and intrinsic homeostasis appear to be linked \cite{Joseph2017}.
-Still, the exact nature of the response depends on how, and to what degree, oscillatory input and other sensory or internally driven ``computational'' inputs share synapses.
-This in turn requires considering complex dendritic arbors and their effect on homeostasis \cite{LeMasson1993}, neuromodulation \cite{Jadi2012,Jadi2014}, and computation \cite{Mainen1996,Polsky2004,Mel2004}.
+We did not study synaptic homeostasis.
+Synaptic homeostasis can also play key a role in combating oscillatory perturbations \cite{Cannon2016}.
+There is some evidence intrinsic homeostasis appear to be linked excitatory synaptic homeostasis \cite{Joseph2017}.
+Still, the exact nature of the homeostatic response depends on how, and to what degree, oscillatory and other sensory or internally driven inputs share synapses.
+This, in turn, requires considering complex dendritic arbors and their effect on homeostasis \cite{LeMasson1993}, neuromodulation \cite{Jadi2012,Jadi2014}, and computation \cite{Mainen1996,Polsky2004,Mel2004}.
 Considering these interacting effects together is the next step needed to develop a clearer biological understanding of modulatory oscillations.
@@
 Here homeostasis has been shown to stabilize self-organized oscillations \cite{Golowasch1999}, and interact with neuromodulation in a \textit{highly} state dependent way \cite{Marder2014,Marder2015,Marder2014}.
-However, this work has focused on how chemical neuromodulators affect the formation of a pacemaker, not how a pacemaker can modulate another, downstream, circuit.
-Which is our focus here.
+Like our model though the oscillation present in the STG is tonic.
+This means it could in principle drive a second homeostatic response.
+However, synapses in the stomatogastric ganglion circuit are predominately inhibitory \cite{Marder2015}.
+If our model predictions held up for this circuit we would expect there to be no second response.
 
-The interaction between homeostasis and oscillations has previously been considered when the oscillatory input is treated as a signal, not a modulator.
+The interaction between homeostasis and oscillations has previously been considered in a simple model of cortex.
+But the oscillatory input here was treated as a signal, not a modulator.
 Cannon and Miller \citep{Cannon2017} explored how synaptic homeostasis can effectively minimize the effect of modulatory perturbations, thus maximizing mutual information between an incoming oscillatory signal and a single cell's firing pattern.
@@
 
- Based on our simple model, we conjecture that intrinsic homeostasis may explain why tonic theta rhythms in the hippocampus are synaptically inhibitory.
+ If our simple model is right, we conjecture that intrinsic homeostasis can explain why tonic theta rhythms in the hippocampus are synaptically inhibitory.
 To make this clear, consider the alternative.
 If the theta rhythm was strong, tonic, and synaptically excitatory, our model suggests this could lead to an equally strong--but opposing--homeostatic response.
-According to our model, such as response means that the firing threshold increases and the likelihood a hippocampal neuron can respond to any given stimulation would decrease, perhaps markedly so.
+Such a response can mean that sodium conductance decreases, which in turn decreases the likelihood a hippocampal neuron could respond to any given stimulation.
 
-In effect, strong excitatory oscillations consume a substantial portion of each cell's possible dynamic range.
-On the other hand inhibitory oscillations do not generate an intrinsic homeostatic response, and so leave the dynamic range of the neurons intact.
-This may also explain why neocortical oscillations tend to be short and bursty, and why some neurological disorders, such as Parkinson's disease, are associated with prolonged rhythms.
+In effect, the homeostatic response to strong excitatory oscillations consumes a substantial portion of each cell's dynamic range.
+On the other hand, inhibitory oscillations--which do not generate an intrinsic homeostatic response--leave the dynamic range of neurons intact.
+This may also explain why neocortical oscillations tend to be short and bursty, and why some neurological disorders, such as Parkinson's disease, are associated with prolonged excitatory rhythms.
 
@@
 \subsection*{Mathematical model}
-We model a feedforward network of hippocampal neurons, subjected to oscillatory modulation.
-This is instantiated as $N = 2000$ input cells connected to $M = 100$ Hodgkin-Huxley neurons.
-The $M$ cells in the network were tuned to mimic regular firing \cite{Borgers2005,Borgers2008}.
+We model an unconnected population of hippocampal neurons, subjected to oscillatory modulation.
+This is instantiated as $N = 1000$ input cells connected to $M = 100$ Hodgkin-Huxley neurons.
+The $M$ cells in the population were tuned to mimic regular firing \cite{Borgers2005,Borgers2008}.
 The firing pattern of each input cell (both stimulus and modulation) is from a Poisson process, with a time-varying rate.
@@
 All input cells have a $p = 0.1$ connection probability to the hippocampal population.
+The hippocampal population has no lateral of recurrent connections.
 The synaptic weights for all $N \rightarrow M$ connections $w$ were independently sampled from a uniform distribution, $w \sim \mathcal{U}(5, 50)$ $\mu$S.
-The firing rate of the oscillating population was governed by sinusoidal pacemaker, with amplitude $A$ and frequency $f$, with the exact form $r_o / 2 (1 + \text{sin}(2 \pi f t)$.
+The firing rate of the oscillating population was governed by sinusoidal pacemaker, with amplitude $A$ and frequency $f$, with the exact form $\frac{r_o}{2} (1 + \text{sin}(2 \pi f t))$.
 As a result, $r_o$ defines the peak firing of the oscillation.
 The stimulus population was simply modeled by a fixed rate of 6 Hz ($r_s$).
-The background firing rate $r_b$ was constant, and set at 2 Hz.
 
@@
 Besides $I_{Ca}$ (which is discussed below), active currents are governed by the standard Hodgkin and Huxley form \cite{Hodgkin1952}.
-Where $m$ and $h$ respectively track the opening and closing channel kinetics, and $E$ is the channel appropriate Nernst reversal potential.
+Where $m$ and $h$ respectively track the opening and closing channel kinetics, $p$ and $q$ are channel dependent parameters, and $V_i$ is the channel appropriate Nernst reversal potential.
 See \textit{Table 1} for the complete set of parameters.
@@
 \label{eq:Idef}
-I = \bar{g} m^p h^q (E - V)
+I = \bar{g} m^p h^q (V_i - V)
 \end{equation}
@@
 
-Overall membrane dynamics were governed by these internal ion conductances, a variable bias current $I_\text{bias}$, and the excitatory synaptic input term $I_S$, which contains both background, stimulus, and oscillatory terms.
+Overall membrane dynamics were governed by these internal ion conductances, a variable bias current $I_\text{bias}$, and the excitatory synaptic input term $I_S$, which contains both stimulus and oscillatory terms.
 All synaptic input was in turn governed a single exponential kinetics, which we denote generically using an $x$ subscript below.
@@
 \begin{equation}
-    I_S = -g_b (E_s - V) + -g_s (E_s - V) + -g_o (E_o - V) 
+    I_S = -g_s (V_e - V) + -g_o (V_o - V) 
 \end{equation}
@@
 \begin{equation}
-C \frac{dV}{dt} = I_l + I_{Na} + I_{K} + I_{KCa} + I_{Ca} + I_{\text{bias}} + I_S
+\label{eq:HHH}
+C \frac{dV}{dt} = I_l + I_{Na} + I_{K} + I_{KCa} + I_{Ca} + I_{\text{bias}} + I_{S} 
 \end{equation}
@@
 During homeostatic equilibration, conductances drifted until the target Ca\textsuperscript{2+} concentration was met, denoted as $C_T$.
-In a control experiment a range of $C_T$ values were explored (Figure \ref{fig:f4}), though simulations defaulting to 0.03 mM; the value the system reaches with stimulation ($r_s = 6$ Hz) without modulation ($r_o = 0$).
+In a control experiment a range of $C_T$ values were explored (Figure \ref{fig:f5}), though simulations default to 0.03 mM; the value the system reaches with stimulation ($r_s = 6$ Hz) without modulation ($r_o = 0$).
 The $\pm$ symbol in equation \ref{eq:dgdt} denotes the direction of ion flow and is $(+)$ for inward going currents (Na and Ca) and $(-)$ for outward going Potassium.
@@
 \begin{equation}
+\label{eq:passive}
 \frac{d\text{Ca}}{dt} = -k \text{Ca} - \gamma I_\text{Ca}
@@
 We measure excitability by comparing average firing rate of all $M$ neurons in an experiment, with ($r_m$) and without modulation ($\bar{r}_m$).
-This accounts for any homeostatic drift driven only by the background noise:
 
@@
 $V_e$ & 0 (mV) & Excitatory synaptic reversal potential \\
-$p$ & 0.1 & Connection probability \\
+$V_i$ & -80 (mV) & Inhibitory synaptic reversal potential \\
+$p$ & 0.1 & Connection probability (stimulus, modulation) \\
 $C_T$ & 0.0028-0.0032 (mM) & Target Ca\textsuperscript{2+} concentration \\

### 2020-03-06 main.tex

@@
 Some oscillations are \textit{tonic} and last for seconds or even minutes.
-Others oscillations are unstable.
-They appear for only as a single-cycle \textit{burst}.
+Other oscillations appear as unstable \textit{bursts}.
 Likewise, some oscillations rely on excitatory AMPAergic synapses, but others are GABAergic and inhibitory.
-Why both kinds of diversity are present is not known.
+Why all this diversity is present isn't well understood.
 We hypothesized Ca$^{2+}$-dependent homeostasis is important to finding an explanation.
@@
 \textsc{\textbf{New and Noteworthy}} The intricate interplay of neuromodulators, like acetylcholine, with homeostasis is well known.
-The interplay between oscillatory modulation and homeostasis is not.
-We studied oscillatory modulation and intrinsic homeostasis for the first time in a simlified model of hippocampal neurons.
-We report a paradoxical result: Ca\textsuperscript{2+}-mediated homeostasis causes normally synchronizing AMPAergic oscillations to become inhibitory and desynchronizing.
-This result, along with other new observations, suggests homeostasis might be just as complex -- and just as important -- for oscillations as it is for all other kinds of modulation.
+The interplay between oscillatory modulation and homeostasis isn't.
+We studied oscillatory modulation and homeostasis for the first time using a simplified model of hippocampus.
+We report a paradoxical result: Ca-mediated homeostasis causes AMPAergic oscillations to become effectively inhibitory.
+This result, along with other new observations, suggests homeostasis might be just as complex and important for oscillations as it for other neuromodulators.
 

### 2020-04-24 response_JNeurophys_2.md
JNeurophys rev 2 - draft 1
- Reviewer text changes are complete
- Revamp Fig 1
- Update text to match this new Fig

# Reviewer #1 (Comments to the Author):

In the revised version of this manuscript, the authors addressed my previous concerns and explained well the details that I asked. By adding the new frequency range to their model, their original claims are more strongly supported. Several clarification troughout the text make the points more clear, specially when comes to the simulation of an artificial oscillation in nature to test their predictions. The discussion makes it more clear now to interpret their results in the context of previous bibliography. Some minor points that I noticed in the text are:

x 1)	Line 144: "Physiological theta is a GABAergic"
x 2)	Line 187: "This means our Ca2+ set point (of 0.003 mM) depends on the gL in practice but our choice of of 0.003 mM was arbitrary
x 3)	Figure 5: Control experiments. A. Change in population hippocampal population rate for different levels of target... B. Change in population synchrony for different levels of target [Ca] (colors) as a function of oscillation...

> We corrected these.

As I previously stated, I find the data in this manuscript very interesting and believe that will be of the general interest of the readers of the journal.

> Again, thank you!

# Reviewer #2 (Comments to the Author):

The manuscript is quite clear now and the results are of interest. Showing that the results can generalize in a number of ways adds to the general importance. I just have a few comments that could be fixed with writing, with no need for new data.

Some confusion could be avoided by for example, omitting all mention of 1000 Poisson neurons and Poisson population (l.67, l. 73) when at the end of the day these neurons are not simulated. Instead an inhomogeneous Poisson spike train is sent to the single neuron being simulated since the sum of 1000 Poisson processes is a single process with 1000 times the rate. I think it is OK to say that you provide such inhomogeneous Poisson input to the single cell and that such input could "represent" the output of 1000 neurons but it is misleading and confusing to suggest that a population of 1000 neurons is included in your model.

e.g. line 77 "instantiation of a randomly generated population" suggests a network with random connections or something. All you did though was instantiated a Poisson process.

> This is not quite right, but your misunderstandings is entirely our fault! We did simulate 1000 independent Poisson cells, but then connected a randomly chosen 10 percent of these to each cell in the HH population. The text around line 77 now reads:

> Specifically, we study Ca\textsuperscript{2+}-mediated homeostasis in a feed-forward population of 100 Hodgkin-Huxley neurons. By feedforward we mean neurons in the population are driven only by the inputs to the model. Cells have no lateral or recurrent connections. We modulate this population with neural oscillations. Oscillatory input for each cell in Hodgkin-Huxley population was modelled as a set of about $N=100$ Poisson neurons. These were randomly drawn from a pool of 1000 Poisson neurons, each of which was independently simulated. To change the level of oscillatory modulation strength, we alter the firing rate of the entire Poisson population. Average synaptic strengths were fixed, but drawn from a uniform distribution for each simulation (see \textit{Methods}). An oscillation's temporal duration was simulated as either tonic, which lasted for the entire experiment, or as a 4-cycle burst. Bursts were delivered only at the end of a trial, where they overlap with the input stimulus (Figure~\ref{fig:f1}\textbf{a}). The stimulus input was taken from an independent set of $N=100$ Poisson cells, that were also drawn from a 1000 cell Poisson population. Though of course stimulus and oscillation populations were independently simulated. The stimulus lasted 0.5 seconds, and had an average firing rate of 6 Hz. The stimulus was fixed for all experiments, and was delivered only once homeostatic equilibrium was reached \cite{Barth2012} (Figure~\ref{fig:f1}). During the stimulus window we measured the population's firing rate and synchrony. It is these numbers we report throughout.

l.94-96 & Fig. 1: Calcium dynamics arise of 100s of milliseconds so there is confusion here. What you mean is the homeostatic response (via signaling pathways including activation of kinases, transcriptional changes, translation of new proteins, movement of mRNA or proteins in the cell and insertion of receptors) to the changes in calcium dynamics which fluctuate on a timescale of a fraction of a second.

> Quite right of course. The text around line 94 now reads:

> Note that in real systems intrinsic homeostasis is thought to happen over minutes or days. However, simulation times that are hours or days long are not computationally feasible. So we follow the field and study a model where Ca\textsuperscript{2+} homeostasis dynamics happen with a 4-second half-life, denoted by $\tau_h$. This might seem like a huge difference. But all that matters mathematically is that homeostasis dynamics happen much slower than all the other synaptic/membrane dynamics. A timescale of $\tau_h > 4$ seconds is a reasonable choice for ensure this. The other major dynamics in the model are synapses with $< 30 ms$ half-lives. This leads to a 133 fold difference between synaptic and homeostatic dynamics \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,Marder2014,OLeary2014,LeMasson1993,Abbott1993}.

l.135 "inversion of kappa": I don't think kappa has been defined before this sentence.

> Corrected. The text around line135 now reads:

> Note that the inversion in population synchrony, seen in Figure~\ref{fig:f2}\textbf{d} as well as in \ref{fig:f5}\textbf{b}, is a low $N$ effect. As population spiking is suppressed, the total number of spikes declines to the point where the bins used to calculate the Kappa correlation often contain no spikes. This leads to a somewhat misleading increase in synchrony.

l. 156-7 "changes due to K ... can't alter baseline". This is unclear or wrong I think. Any change in K-conductance in a Hodgkin-Huxley like model (like changing the leak conductance) alters both mean firing rate and mean membrane potential depending what is meant by baseline. (A simple simulation reveals this).

> Recovery and initiation are entangled, of course. The text around line 156 now reads:

> The excitability changes we see with homeostasis must be due to the decreases in Na conductance (which initiates spikes) and increases in K conductance (which recovers the voltage following a spike).

Overall I think some more intuition could be provided into why the main result arises, again, more is needed than a simulation demonstrating an effect. Rather *a clear explanation*.

> TODO - I have no idea how to state it more clearly... Write more random text?

Typos:
x l.23 missing "is"
x l.61 "suppresses" is wrong
x l.105 "the oscillation's"
x l.144 extra "a"
x l.158 missing "of"
x l. 177 missing "long" after "how"

> Thank you.

# Reviewer #3 (Comments to the Author):

Thank you for the improvement over the original submitted manuscript, which has helped me fill in some of the blanks that were previously missing. This is an interesting exploration to consider how homeostatic mechanisms would differently affect AMPAergic versus GABAergic oscillatory inputs. However, there is still room for improvement and clarification to help the reader appreciate and understand details of the work done.

1) Please help the reader by specifying all aspects clearly in the model set-up, equation parameters and interpretation used. Specifically:

(i) I_S referred to only as excitatory in methods (eq 3), but both AMPA and GABA -ergic were examined? It is not clear, largely because parameters are not fully described I think. For example, does g_o, V_o refer to either excitatory or inhibitory and so V_o=V_i or V_e? what is g_x refer to, presumably either but there seem to be too many parameters introduced? I'm guessing the tau_b etc. are excitatory or inhibitory time constants? Could the authors simply fully write out the equations for excitatory or inhibitory oscillatory modulation separately to be clear. Also, parameter values for I_bias (eq 5) are missing - was it varied in the 100 HH models? I'm guessing not as it would add to the variability and synchrony of output to complicate interpretation?

> Point (i) made a few points. Point by point:

> We now better explain all synaptic dynamics.
> The text "I_S referred to only as excitatory in methods" was corrected. It now reads, "Overall membrane dynamics were governed by these internal ion conductances, and the synaptic input term $I_s$, which contains both stimulus and oscillatory terms."
> I_bias was set to zero in all simulations. So we removed it from the paper.

(ii) Could the authors expand/improve Fig 1 to help the reader please. In particular, the 'model input' is a bit confusing to me - as I understand it, there are 2 different types of input added at the last 0.5 sec - tonic or bursting oscillation. As presented, one might think this is the top and bottom rows, but it is the 2nd and 3rd rows as given in the caption. So, for the tonic input, there is nothing specifically changed during the last 0.5 sec? That is, it is just an ongoing oscillatory input? I wonder if the depiction is due to different I_bias values? I hope the authors can make this clear to the reader by fully describing these model input depictions. Model input is a raster plot, right? Expansions are to show that there are no population bursts (top) but there are in 2nd and 3rd panels (?) etc. For the model output in Fig.1c, it would be helpful to also see an expansion of the last 0.5 sec too for the HH cell, perhaps showing an example for the tonic and bursting output.

(iii) With the above point clear, the authors can now provide the reader with a clear description of Fig. 2 and onward regarding oscillation modulation (values of 0-3). As they specify in the revision, this is due to the altering the firing rate of the Poisson population (fixed rate of 6 Hz was changed in the modulation there, to what? etc.?). Perhaps this can be illustrated (akin to revised Fig 1a aspects). I assume that these modulation values are arbitrary values? The authors could specify this clearly in the Methods and/or expanded caption. That is, please 'map' what parameter values in equations to what is specifically plotted. As I understand it from the methods, half of the 1000 Poisson cells have oscillatory aspects (N_o) and the others don't, and presumably they are not connected themselves, right? Is the amplitude A meant to be r_o/2? Not sure why parameter 'A' needs to be introduced?

> TODO parameters for example plots in Fig 1? Because, sigh, seriously lady oh my god.

> 'A' was an arifact from a previous draft. We removed it. It should be, as you say, r_0/2.

2) Fig 4 shows average - what was the variance like? Presumably if this was done with inhibitory oscillatory input, it would not change much from the 'no homeostasis' FI curve? It would be helpful to show this to make their paradoxical result clear.

3) It might be interesting to note rodent and human theta hippocampal oscillation differences (e.g., see J.Jacobs Phil Transc. 2014 where intermittent human oscillations described)

Minor:
x 1) line 52 - "Given that chemical..."
x 2) line 61 - "suppress..."2) line 61 - "suppress..."
x 3) lines 133-134 - "...so most probably incomplete at best. Experimental testing is needed". As the authors state, model is 'very simple', one can simply say that it "is" incomplete. What sort of experimental testing do the authors envision?

> We updated the text to both fix the typo, and we provided a sketch for testing our work in vivo. It now reads,

> The general effect we would expect for some other selection of channels, chosen by some other theorists, or by evolution itself, depends on those selected channels. O'Leary's \cite{OLeary2014} has made is clear how complex the relationship is between homeostasis, channel conductance, and firing patterns. It's so complex that we believe general statements are probably impossible. On the one hand then, our simple model has been shown by others to recapitulate hippocampal pyramidal cell's activity in laboratory settings \cite{LeMasson1993}. On the other, it is \textit{very simple} and validating our work requires experimental testing. As an example, by using prolonged excitatory gamma oscillations--arising from direct stimulation of either the dentate gyrus or select areas in CA1 \cite{Csicsvari2003}--to drive a possible homeostatic response in CA3.

x 4) lines 144-5 - not a sentence "Physiological theta....".

Please add some references for this, e.g., Colgin. In their simple model, the authors are presumably assuming that the theta rhythm is primarily due to external oscillatory input (medial septum), but it is of course now clear that there are intra-hippocampal theta rhythms (Williams group, and as notes in Colgin reviews).

> We specified the generator of interest in our work, including the requested citation. The text now read:

> In Figure~\ref{fig:f2} we compared AMPA and GABAergic models of the theta rhythm (8 Hz). This was done to illuminate their different effects, but was not a biologically realistic model. Hippocampal theta is solely a GABAergic rhythm, at least when driven by medial septum \cite{Colgin2016a}. In hippocampus AMPAergic oscillations are, however, seen in the high gamma range (60-90 Hz). In Figure~\ref{fig:f3} we confirm AMPAergic oscillations in the gamma range follow the same pattern as theta. And they do. For both 60 and 90 Hz gamma, tonic oscillations suppress excitability and desynchronize firing.

x 5) line 250 - " ...evidence that intrinsic..."
x 6) lines 263-4 - STG and acronym used and then not on these 2 lines - please fix.
x 7) line 299 - "... by a sinusoidal..."

As I would prefer reviews to not be anonymous, I would like to identify myself (Frances Skinner).

### 2020-04-24 main.tex
JNeurophys rev 2 - draft 1
- Reviewer text changes are complete
- Revamp Fig 1
- Update text to match this new Fig

@@
 \begin{abstract}
-Neural oscillations are observed ubiquitously in the mammalian brain, but their stability is quite variable.
+Neural oscillations are observed ubiquitously in the mammalian brain, but their stability is known to be rather variable.
 Some oscillations are \textit{tonic} and last for seconds or even minutes.
@@
 Likewise, some oscillations rely on excitatory AMPAergic synapses, but others are GABAergic and inhibitory.
-Why all this diversity is present isn't well understood.
-We hypothesized Ca$^{2+}$-dependent homeostasis is important to finding an explanation.
-We tested this in a highly simplified model of hippocampal neurons.
-In this model homeostasis can profoundly alter the modulatory effect of neural oscillations.
-Under homeostasis, tonic AMPAergic oscillations actually \textit{decrease} excitability and desynchronize firing.
+Why this diversity exists is not clear.
+We hypothesized Ca$^{2+}$-dependent homeostasis could be important in finding an explanation.
+We tested this hypothesis in a highly simplified model of hippocampal neurons.
+In this model homeostasis profoundly alters the modulatory effect of neural oscillations.
+Under homeostasis, tonic AMPAergic oscillations actually \textit{decrease} excitability and \textit{desynchronize} firing.
 Tonic oscillations that are synaptically GABAergic--like those in real hippocampus--don't provoke a homeostatic response, however.
-If our simple model is correct, homeostasis can explain why the theta rhythm in hippocampus is synaptically inhibitory: GABA has little to no intrinsic homeostatic response, and so can preserve a cell's dynamic range under tonic oscillation.
-Based on these results we also speculate that homeostasis can explain why AMPAergic oscillations in cortex often appear as bursts.
-Bursts minimally interact with the slow homeostasis time constant, and so retain their excitatory effect.\\
+If our simple model is correct, homeostasis can explain why the theta rhythm in hippocampus is synaptically inhibitory: GABA has little to no intrinsic homeostatic response, and so can preserve the pyramidal cell's natural dynamic range.
+Based on these results we can also speculate that homeostasis may explain why AMPAergic oscillations in cortex, and hippocampus, often appear as bursts.
+Bursts do not interact with the slow homeostatic time constant, and so retain they their normal excitatory effect.\\
 \end{abstract}
@@
 \textsc{\textbf{New and Noteworthy}} The intricate interplay of neuromodulators, like acetylcholine, with homeostasis is well known.
-The interplay between oscillatory modulation and homeostasis isn't.
+The interplay between oscillatory modulation and homeostasis is not.
 We studied oscillatory modulation and homeostasis for the first time using a simplified model of hippocampus.
 We report a paradoxical result: Ca-mediated homeostasis causes AMPAergic oscillations to become effectively inhibitory.
-This result, along with other new observations, suggests homeostasis might be just as complex and important for oscillations as it for other neuromodulators.
+This result, along with other new observations, means homeostasis might be just as complex and important for oscillations as it is for other neuromodulators.
 
 \section*{Introduction}
-\lettrine[loversize=0,nindent=0,realheight=true]{N}{} euromodulation and homeostasis are inter-linked but opposing phenomena.
-Modulation perturbs excitability, which we define as the propensity for a stimulus to elicit an action potential.
+\lettrine[loversize=0,nindent=0,realheight=true]{N}{}euromodulation and homeostasis can be thought of as inter-linked but opposing phenomena.
+Neuromodulation perturbs excitability, which we define as the propensity for a stimulus to elicit an action potential.
 Homeostasis acts to quench this perturbation, driving the excitability of the cell back to a biologically desirable set point \cite{LeMasson1993,Abbott1993}.
-While the interplay between chemical modulators and homeostasis has been studied for more than 20 years \cite{LeMasson1993,Abbott1993,Golowasch1999,Marder2014,Marder2015,Gutierrez2013}, the relationship between population-level synaptic modulations--like neural oscillations--and homeostasis is not well understood, on either theoretical or empirical grounds.
-Like neuromodulators, neural oscillations alter excitability and firing statistics.
-Uniquely though, oscillations can group action potentials into synchronous windows of activity \cite{Lisman2013,Voytek2015}.
-Grouping action potentials can improve signal to noise and increase the number of coincident firing events \cite{Chen2013,Zhou2015,Voytek2015a,Peterson2017}, which drives learning at individual synapses \cite{Muller2011,Song2000,Markram1997}.
+While the interplay between chemical modulators and homeostasis has been studied for more than 20 years \cite{LeMasson1993,Abbott1993,Golowasch1999,Marder2014,Marder2015,Gutierrez2013}, the relationship between population-level synaptic modulations--like neural oscillations--and homeostasis is not well understood.
+It is not well understood theoretical, or well-studied experimentally.
 
-We propose the same mechanisms that link homeostasis with chemical neuromodulation can also come into play with oscillatory modulation.
-After all, both kinds of modulation lead to tonic changes in spiking and, as a result, changes in Ca\textsuperscript{2+} concentration \cite{Liu1998}.
+Like neuromodulators, neural oscillations are ubiquitous and alter the excitability and firing statistics of cells throughout the brain.
+Uniquely though, oscillations group action potentials into synchronous windows of activity \cite{Lisman2013,Voytek2015}.
+Grouping action potentials improves their signal to noise, increases the number of coincident firing events \cite{Chen2013,Zhou2015,Voytek2015a,Peterson2017}, that in turn drives learning at individual synapses \cite{Muller2011,Song2000,Markram1997}.
 
+We propose the same mechanisms that link homeostasis with chemical neuromodulation should also come into play with oscillatory modulation.
+After all, both kinds of modulation can lead to tonic changes in spiking and, as a result, changes in Ca\textsuperscript{2+} concentration \cite{Liu1998}.
 To test this, we model activity-dependent intrinsic homeostasis in a feed-forward population of hippocampal pyramidal cells \cite{Siegel1994}.
 By feed-forward we mean there are no lateral or recurrent connections in the population.
-Firing activity is driven only by afferent synaptic input, and homeostasis is mediated by a Ca\textsuperscript{2+}-dependent mechanism \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,OLeary2014} pioneered by LeMasson \cite{LeMasson1993,Abbott1993}.
-Here Ca\textsuperscript{2+} acts a sensor for tonic changes in the membrane voltage.
+Firing activity is driven only by afferent synaptic input, and homeostasis is mediated by a Ca\textsuperscript{2+}-dependent mechanism \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,OLeary2014} whose early work was pioneered by LeMasson \cite{LeMasson1993,Abbott1993}.
+In our model Ca\textsuperscript{2+} acts a sensor for tonic changes in the membrane voltage.
 To counter tonic changes in Ca\textsuperscript{2+} levels, the expression of ion channels is altered, returning the Ca\textsuperscript{2+} level to a predefined ``good'' value \cite{Golowasch1999,OLeary2013}.
-Following Siegel \cite{Siegel1994}, increases in Ca\textsuperscript{2+} lead to downregulation of Na\textsuperscript{+} and Ca\textsuperscript{2+} channels, and upregulation of fast K\textsuperscript{+} channels.
-To minimize the effect of homeostasis on the rise and fall of action potential dynamics, we also add a KCa channel which is not present in Siegel \citep{Siegel1994}.
+Following Siegel \cite{Siegel1994}, increases in Ca\textsuperscript{2+} lead to downregulation of Na\textsuperscript{+} and upregulation of fast K\textsuperscript{+} channels.
+To minimize the effect of homeostasis on overall action potential magnitude, we also added a KCa channel \citep{Siegel1994}.
 In real cells intrinsic homeostasis changes the expression level of ion channels \cite{OLeary2013}.
@@
 
-Given chemical modulators operate on both short and long times scales \cite{Marinelli2014,Marder2014,Cohen2015,Daw2002}, we examined two timescales of oscillatory modulation.
-We contrast the effects of short 4-cycle bursts of oscillation to long lasting tonic rhythms.
+Given that chemical modulators operate on both short and long times scales \cite{Marinelli2014,Marder2014,Cohen2015,Daw2002}, we examined two timescales of oscillatory modulation.
+We contrast the effects of short 3-cycle bursts of oscillation to long lasting tonic rhythms.
 Both are observed {\textit{in vivo}}, and are known to play different physiological, cognitive, and computational roles \cite{Lundqvist2016,vanEde2018,Peterson2017}.
@@
 \section*{Results}
-We assessed the impact of homeostasis in a simple model that received oscillatory inputs that are either transiently bursting, or tonic.
-We found that increasing the strength of AMPAergic oscillatory inputs resulted in a homeostatic response that, when reaching an appropriate level of average calcium, overcompensates to actually suppresses the postsynaptic neuron's spiking.
-That is, paradoxically, an increase in drive when combined with homeostasis results in a decreased response.
+We assessed the impact of homeostasis in a simple model that received oscillatory inputs that are either transiently bursting, or long lasting and tonic.
+We found that increasing the strength of AMPAergic oscillatory inputs resulted in a homeostatic response that, when reaching an appropriate level of average calcium, overcompensates to actually decrease the postsynaptic neuron's spiking.
+That is, paradoxically, an increase in drive-- when combined with homeostasis--results in a decreased response.
 
@@
 We modulate this population with neural oscillations.
-Oscillations are modelled as a set of 1000 Poisson neurons.
-To change the level of oscillatory modulation strength, we alter the firing rate of the Poisson population.
+Oscillatory input for each cell in Hodgkin-Huxley population was modelled as a set of about $M=100$ Poisson neurons (Figure~\ref{fig:f1}\textbf{b}-\textbf{c}).
+With a probability of 0.1 these were randomly drawn from a pool of 1000 Poisson neurons, each of which was independently simulated.
+To change the level of oscillatory modulation strength, we alter the firing rate of the entire Poisson population.
 Average synaptic strengths were fixed, but drawn from a uniform distribution for each simulation (see \textit{Methods}).
-An oscillation's temporal duration was simulated as either tonic, which lasted for the entire experiment, or as a 4-cycle burst.
-Bursts were delivered only at the end of a trial, where they overlap with the input stimulus (Figure~\ref{fig:f1}\textbf{a}).
-The stimulus input was also drawn from a 1000 cell Poisson population.
-It lasted 0.5 seconds, and had an average firing rate of 6 Hz.
-The stimulus was fixed for all experiments, and was delivered only once homeostatic equilibrium was reached \cite{Barth2012} (Figure~\ref{fig:f1}).
+An oscillation's temporal duration was simulated as either tonic (Figure~\ref{fig:f1}\textbf{b}, top), which lasted for the entire experiment (20 sec), or as a 3-cycle burst (Figure~\ref{fig:f1}\textbf{b}, bottom).
+Bursts were delivered only at the end of a trial, where they overlap with the input stimulus (Figure~\ref{fig:f1}\textbf{a}-\textbf{b}).
+The stimulus input was also taken from an independent set of about $M=100$ Poisson cells.
+Stimulus and oscillation populations were independently simulated.
+The stimulus lasted 0.5 seconds, and had an average firing rate of 6 Hz, and was fixed for all experiments.
 During the stimulus window we measured the population's firing rate and synchrony.
@@
 Each experiment began with an instantiation of a randomly generated population.
-A single example neuron is depicted in Figure~\ref{fig:f1}\textbf{b}.
+A single example neuron is depicted in Figure~\ref{fig:f1}\textbf{c}.
 This population is subjected to a range of modulatory and control conditions, including oscillatory strength, duration, and synapse type (AMPA or GABA).
 Each experiment lasted 20 seconds.
-In Figure~\ref{fig:f1}\textbf{c} we depict key aspects of model output during an experiment.
+In Figure~\ref{fig:f1}\textbf{d}-\textbf{g} we depict key aspects of model output during an experiment.
 
@@
 Both of these measures are defined in the \textit{Methods}.
-To ensure a consistent comparison between experiments, measurements were made over the same 4-cycle or 0.5 second period in all simulations.
+To ensure a consistent comparison between experiments, measurements were made over the same 0.5 second period in all simulations.
 Source code for all simulations is available at \url{https://github.com/voytekresearch/resistingrhythm}
 
-The homeostatic response in our simple model decreases the sodium conductance and increases the potassium conductance (Figure~\ref{fig:f1}\textbf{c}, \textit{bottom panel}).
-Combined these make it harder for the model cell to generate action potentials.
-This basic mechanism explains the loss of excitability in some conditions of our model and, in turn, the desynchronization of population activity.
+The homeostatic response in our simple model decreases the sodium conductance and increases potassium conductances (Figure~\ref{fig:f1}\textbf{g}, \textit{bottom panel}).
+Combined, these make it harder for the model cell to generate action potentials and extends the recovery period after a spike.
+These two changes combine to explain the loss of excitability in some conditions of our model.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{fig1.png}
+\includegraphics[width=1\textwidth]{fig1.png}
 \caption{\label{fig:f1}
-    Diagram of the model.
-\textbf{a}.
+    The model, its inputs, and some example results. We show examples with and without homeostasis, of both bursts (goldenrod) and tonic oscillations (black).
+    \textbf{a}-\textbf{b}.
 Inputs into the model.
-Top panel depicts the stimulus, applied over the last 0.5 seconds of each simulated trial.
-Middle and bottom panels depict the two modes of oscillation we examined: tonic and bursting.
-    \textbf{b}. Illustration of a single model neuron, its major currents, its inputs (top arrow) and its output (bottom arrow).
-    \textbf{c.} Examples of model output, including membrane potential (top panel), Ca\textsuperscript{2+} concentration, both observed (solid) and the homeostatic target (dotted line).
-The bottom panel in c.
-shows homeostatic dynamics beginning with trial onset, followed by a delay to equilibrium over 20 seconds.
-Note that there is no homeostatic response to stimulus onset at 19.5 seconds.
-Also note the log scale on the y axis in this panel.
+\textbf{a.} The stimulus, which is used to estimate the model population's synchrony and excitability.
+\textbf{b.} Two example input oscillations, a tonic example (black) and a burst example (goldenrod).
+    \textbf{c.} Illustration of a single model neuron, its major currents, and the major equations for intrinsic homeostasis. 
+    \textit{Note}: the insets in for the panels above depict input spikes over the last 1 second of run time.
+    \textbf{d.}-\textbf{e.} Example of model output with homeostasis dynamics turned off.
+\textbf{d.} Is the membrane voltage, plotted in time.
+\textbf{e.} Depicts the Na, K, and KCa conductances, which are in this case fixed.
+    \textbf{f.}-\textbf{g.} Example of model output, with the homeostasis dynamics turned on.
+\textbf{f.} Is the membrane voltage, plotted in time.
+\textbf{g.} Depicts changes to Na, K, and KCa conductances with time.
 }
@@
 However, simulation times that are hours or days long are not computationally feasible.
-So we follow the field and study a model where Ca\textsuperscript{2+} dynamics happen with a 4-second half-life, denoted by $\tau_h$.
+So we follow the field and study a model where Ca\textsuperscript{2+} homeostasis dynamics happen with a 4-second half-life, denoted by $\tau_h$.
 This might seem like a huge difference.
-But all that matters mathematically is that Ca\textsuperscript{2+} dynamics happen much slower than all the other synaptic/membrane dynamics.
+But all that matters mathematically is that homeostasis dynamics happen much slower than all the other synaptic/membrane dynamics.
 A timescale of $\tau_h > 4$ seconds is a reasonable choice to ensure this.
-The other major dynamics in the model are synapses with $< 30 ms$ half-lives.
-This leads to a 133 fold difference \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,Marder2014,OLeary2014,LeMasson1993,Abbott1993}.
+The other major dynamics in the model are synapses with $< 30$ ms half-lives, leaving us with a 133-fold difference between the time scale of synaptic and homeostatic dynamics \cite{Golowasch1999,Marder2014,Marder2015,Gutierrez2013,Marder2014,OLeary2014,LeMasson1993,Abbott1993}.
 
@@
 Homeostasis completely inverts the effect of tonic AMPAergic oscillatory modulation.
-These excitatory oscillations now result in decreases in both population firing rate, and population synchrony.
-To see how, first consider how the population responds to a tonic excitatory modulation \textit{without} homeostasis.
-Without homeostasis, increasing the strength of the excitatory oscillation increases excitability and increases the population firing rate.
-The action potentials are also grouped by oscillation's rising phase, increasing the population's synchrony (Figure \ref{fig:f2}\textbf{b}, black line).
+These excitatory oscillations cause decreases in population firing rate and population synchrony.
+To see how, let's consider how the population responds to a tonic excitatory modulation \textit{without} homeostasis.
+Without homeostasis, increasing the strength of the excitatory oscillation increases excitability, and so increases the population firing rate (Figure \ref{fig:f2}\textbf{b}, black line).
 
-With homeostasis this pattern inverts.
-As excitatory oscillatory modulation strength increases, synchrony and excitability \emph{decreases} (Figure \ref{fig:f2}\textbf{c-d}).
+With homeostasis in play, this pattern inverts.
+As excitatory oscillatory modulation strength increases, synchrony and excitability \emph{decrease} (Figure \ref{fig:f2}\textbf{c-d}).
 Homeostatic mechanisms in the model cause what \emph{should be} excitatory synchronizing modulation to become suppressive and desynchronizing.
 The stronger the oscillation, the more suppressive the result.
-By the time the oscillation is about half the strength of the stimulus (which we fix at a firing rate of 6 Hz) the stimulus is completely suppressed and the population firing rate approaches 0 (Figure \ref{fig:f2}\textbf{c}).
+By the time the oscillation is about half the strength of the stimulus (which we fix at a firing rate of 6 Hz) the response to the stimulus is completely suppressed and the population firing rate approaches 0 (Figure \ref{fig:f2}\textbf{c}).
 
-We compared the effect of tonic oscillation to short 4-cycle bursts of excitatory modulation, presented only during the stimulus.
-Here, the oscillation period is far too short to provoke homeostasis.
-This means that increases in oscillatory strength continue to increase population firing rate and synchrony (Figure~\ref{fig:f2}\textbf{e} and \textbf{f}).
+We compared the effect of these tonic oscillations to short 3-cycle bursts of modulation, presented only during the stimulus.
+Here, the oscillation period is far too short to provoke much, if any, homeostasis.
+This lets bursts sidestep our early results, and so increases in oscillatory strength continue to increase population firing rate and synchrony (Figure~\ref{fig:f2}\textbf{e} and \textbf{f}).
 
-Of course theta oscillations in hippocampus are GABAergic, not AMPAergic as we model here.
-Exploring this artificial oscillation lets us show the central principle we want to highlight in our simple model.
-It seems that tonic oscillations--when joined with homeostasis--can profoundly alter the coding properties of what one would expect from oscillatory modulation.
-The degree of effect depends on the degree to which the oscillation changes the cell's tonic Ca$^{2+}$.
-It also shows that because of homeostasis, bursts are \textit{qualitatively distinct} from their tonic counterparts.
+Of course theta oscillations in hippocampus are GABAergic \cite{Colgin2016}, not AMPAergic as we model here.
+However by exploring artificial AMPAergic oscillation we can demonstrate the central principle we want to highlight.
+It seems that tonic oscillations--when joined with homeostasis--can profoundly alter the coding properties of what one would typically expect from oscillatory modulation, and the degree of this effect depends entirely on the degree to which the oscillation changes the cell's tonic Ca$^{2+}$.
+And that because of this, bursts of oscillation are \textit{qualitatively distinct} from their tonic counterparts.
 
-The specific effect we observed (suppression of excitability) is seen because the homeostatic equations that \textit{we chose} respond to changes in Ca\textsuperscript{2+} by decreasing the conductance of the Na and Ca channels, and increasing the conductance of K and KCa channels (Eq~\ref{eq:dgdt}).
-The net effect of these dynamics is to decrease the sodium channel conductance, and increase the potassium.
+The specific effect we observed (suppression of excitability) is seen because the homeostatic equations that \textit{we chose} respond to changes in Ca\textsuperscript{2+} by decreasing the conductance of the Na, and increasing the conductance of K and KCa channels (Eq~\ref{eq:dgdt}).
 This makes firing an action potential less likely, for any given input.
-These channel dynamics are visualized in Figure~\ref{fig:f1}\textbf{c} and the corresponding FI curve is shown in Figure~\ref{fig:f4}.
+These channel dynamics are visualized in Figure~\ref{fig:f1}\textbf{g} and a corresponding FI curve is shown in Figure~\ref{fig:f4}.
 
-The general effect we would expect for some other selection of channels, chosen by some other theorists or by evolution itself, depends on those channels.
-O'Leary \cite{OLeary2014} has shown how complex the relationship is between homeostasis, channel conductance, and neuronal firing patterns.
-It's complex enough that general statements are probably impossible.
-On the one hand, our simple model has been shown by others to recapitulate hippocampal pyramidal cell's activity in laboratory settings \cite{LeMasson1993}.
-On the other, it is \textit{very simple} and so most probably incomplete at best.
-Experimental testing is needed.
+The general effect we would expect for some other selection of channels, chosen by some other theorists, or by evolution itself, depends strongly on those selected channels.
+O'Leary \cite{OLeary2014} has made it clear how complex the relationship is between homeostasis, channel conductance, and firing patterns.
+It's so complex that we believe general statements are probably impossible, at least as they might relate to this paper.
+While our simple model has been shown by others to recapitulate hippocampal pyramidal cell's activity in laboratory settings \cite{LeMasson1993}, it is a \textit{very simple} model, and validating our work does require new experimental testing.
+This could be done, for example, by using prolonged excitatory gamma oscillations--arising from direct stimulation of either the dentate gyrus or select areas in CA3 \cite{Csicsvari2003}--to drive a possible homeostatic response in CA1.
 
-Note that the inversion in Kappa seen in Figure~\ref{fig:f2}\textbf{d} as well as in \ref{fig:f5}\textbf{b}, is the result of the decrease in excitability.
-Put another way, it's a low $N$ effect.
-As population spiking is suppressed, the total number of spikes declines to the point where the bins used to calculate the Kappa correlation often contain no spikes.
-This leads to a somewhat misleading increase in synchrony.
+Our model shows that there is sometimes an inversion in population synchrony, as seen in Figure~\ref{fig:f2}\textbf{d} and \ref{fig:f5}\textbf{b}.
+We believe this is a low $N$ effect: as population spiking is suppressed, the total number of spikes declines to the point where the bins used to calculate the Kappa correlation are often empty.
+This leads to a somewhat misleading ``increase'' in synchrony.
 
@@
 This is true for both tonic oscillations and bursts.
-In all cases, as oscillatory strength increases, population firing declines dramatically (grey lines in Figure \ref{fig:f2}\textbf{a-d} and light yellow lines in panel \textbf{e-f}).
+No matter the homeostatic state, as GABAergic strength increases population firing declines (grey lines in Figure \ref{fig:f2}\textbf{a-d} and light yellow lines in panel \textbf{e-f}).
 
@@
 
-In Figure~\ref{fig:f2} we compared AMPA and GABAergic models of the theta rhythm (8 Hz).
-This is done to illuminate their differences, but is not biologically realistic.
-Physiological theta is a GABAergic.
-AMPAergic oscillations of the hippocampus are, however, seen in the high gamma range (60-90 Hz).
-In Figure~\ref{fig:f3} we confirm AMPAergic oscillations in this range follow the same pattern as theta.
+In the work so far, Figure~\ref{fig:f2}, we compared AMPA and GABAergic models of the theta rhythm (8 Hz).
+This was done to illuminate their different effects, but was not a biologically realistic model.
+Hippocampal theta is solely a GABAergic rhythm, at least when driven by medial septum \cite{Colgin2016a}.
+AMPAergic oscillations are, however, seen in the high gamma range (60-90 Hz) \cite{Csicsvari2003,Colgin2016,}.
+In Figure~\ref{fig:f3} we confirm AMPAergic oscillations in the gamma range follow the same pattern as theta.
+And they do.
 For both 60 and 90 Hz gamma, tonic oscillations suppress excitability and desynchronize firing.
@@
 \caption{\label{fig:f3}
-    The effect of fast AMPAergic gamma oscillations for both tonic oscillation and 4-cycle bursts. 
+    The effect of fast AMPAergic gamma oscillations for both tonic oscillation and 3-cycle bursts. 
     Top panel (\textbf{a}.) is the observed population firing rate averaged over a 0.05 second or 0.033 second stimulus.
@@
 \subsection*{Homeostasis and excitability.}
-In our model Ca\textsuperscript{2+} influx is driven by increases in the membrane voltage, and Ca\textsuperscript{2+} concentration is restored both by a passive and linear recovery (Eq~\ref{eq:passive}) and active homeostasis (Eq.~\ref{eq:dgdt}).
+In our model Ca\textsuperscript{2+} influx is driven by increases in the membrane voltage but the Ca\textsuperscript{2+} concentration is restored by passive and linear recovery (Eq~\ref{eq:passive}), and active homeostasis (Eq.~\ref{eq:dgdt}).
 Homeostasis works by negative feedback.
-Increases in Ca\textsuperscript{2+} concentration decrease the conductance of both the sodium and potassium channels in our model (Eq.~\ref{eq:HHH}).
+Increases in Ca\textsuperscript{2+} concentration decrease the conductance of both the Na and K channels in our model (Eq.~\ref{eq:HHH}).
 This decrease in conductance in turn lowers the membrane voltage down closer to its resting value.
 As the voltage falls so does the Ca\textsuperscript{2+} concentration.
-The excitability changes we see with homeostasis must be due to the decrease in Na conductance.
-Changes due to K conductance--which acts to recover the voltage following a spike--can't alter baseline.
+The excitability changes we see with homeostasis must be due to the decreases in Na conductance (which initiates spikes) and increases in K conductance (which recovers the voltage following a spike).
 
-One the simplest, classic ways to measure the excitability of a cell--simulated or real--is with an FI-curve.
+One the simplest, most classic ways to measure the excitability of a cell--simulated or real--is with an FI-curve.
 In an FI-curve a small square wave of current is injected into the cell, and the resulting increase in firing rate (if any) is averaged over a short window.
@@
 
-We chose a 4-cycle length for bursts somewhat arbitrarily.
+We chose a 3-cycle length for bursts somewhat arbitrarily.
 Real oscillations range in duration from single cycles to seconds of even minutes \cite{Lundqvist2016,vanEde2018}.
 Our burst/tonic distinction is therefore a false (but useful) dichotomy.
-To find a better estimate of how an oscillation would need to last, on average, to generate a homeostatic response we ran some long simulations using a more realistic homeostatic time constant ($\tau_h$ = 600 seconds; up from 20 seconds).
+To find a better estimate of how long an oscillation would need to last, on average, to generate a homeostatic response we ran some long simulations using a more realistic homeostatic time constant ($\tau_h$ = 600 seconds; up from 20 seconds).
 It seems an oscillation would need to be present \textit{at least} 10\% of the time to generate a response (Figure~\ref{fig:f5}\textbf{c}).
-
 Real theta rhythms drift in frequency \cite{Buzsaki2015}.
@@
 The leak conductance ($g_L$, Eq.~\ref{eq:HHH}) plays a key role in setting the resting potential of the cell.
-The resting potential in turn helps determines overall membrane excitability, and sets the resting Ca\textsuperscript{2+} concentration.
-This means our Ca\textsuperscript{2+} set point (of 0.003 mM) depends on the $g_L$ in practice but our choice of of 0.003 mM was arbitrary.
-To test if a small mis-match between $g_L$ and the set-point altered our results, we ran several experiments perturbing the leak conductance while keeping the Ca\textsuperscript{2+} set-point fixed.
-Changes to leak conductance had only a small effect on homeostasis (Figure~\ref{fig:f5}\textbf{f}-\textbf{g}).
+The resting potential in turn helps determine the overall membrane excitability, and sets the resting Ca\textsuperscript{2+} concentration.
+This means our Ca\textsuperscript{2+} set point (of 0.003 mM) depends on $g_L$.
+To test if a small mismatch between $g_L$ and the Ca\textsuperscript{2+} set-point altered our results, we ran several experiments perturbing the leak conductance while keeping Ca\textsuperscript{2+} fixed, but in the end mismatch in leak conductance had only a small effect on homeostasis (Figure~\ref{fig:f5}\textbf{f}-\textbf{g}).
 
@@
     Control experiments.
-    \textbf{a.} Change in population hippocampal population rate for different levels of target [Ca] (colors) as a function oscillation strength (Osc.
-rate on the x-axis).
-All values are referenced to a no-modulation control.
-    \textbf{b.} Change in population synchrony for different levels of target [Ca] (colors) as a function oscillation strength (firing rate). All values are referenced to a no-modulation control.
+    \textbf{a.} Population firing rate for different levels of target [Ca] (colors). All values are referenced to a no-modulation control.
+    \textbf{b.} Change in population synchrony for different levels of target [Ca] (colors). All values are referenced to a no-modulation control.
     \textbf{c.} Change in population rate as the oscillation duration approaches a more realistic $\tau_h$, the half-life of the homeostasis dynamics (Eq.~\ref{eq:dgdt}).
@@
 Our scientific understanding of homeostasis has been shaped as much by theoretical work as empirical \cite{Marder2014}.
-In an attempt to understand the interaction between oscillatory modulations and homeostasis, we begin by studying one of the simplest models used in early studies of homeostasis--a population of point neurons in the hippocampus \cite{LeMasson1993}.
+In an attempt to understand the interaction between oscillatory modulations and homeostasis, we began by studying one of the simplest models used in early studies of homeostasis--a population of point neurons in the hippocampus \cite{LeMasson1993}.
 
@@
 
+Oscillations are generally conceived to play a functional role in modulating spiking {Cite Varela}{Cite Fries}, by grouping action potentials and improving signal to noise \cite{Chen2013,Zhou2015,Voytek2015a,Peterson2017}.
+Absent the effects of homeostasis, our model clearly demonstrates this: with increasing modulation strength, excitatory oscillations show increased firing rate and synchrony.
+Given this effect, and the importance of phase coding in the hippocampus {Cite}, it is reasonable to wonder why hippocampal oscillations are GABAergic.
+Our model suggests that while this is true for brief, bursty excitatory oscillations, stronger, tonic excitatory oscillations engage homeostatic mechanisms that alter Na and K conductances, driving down the membrane voltage.
+This, in turn, reduces both the firing rate and spike synchrony in response to inputs.
+
 Homeostasis is generally viewed as a stabilizing force.
 This is the goal of homeostasis in our model as well.
-However oscillatory modulation is a unique form of modulation; It shares neurotransmitters, and even synapses, with non-modulatory stimuli.
+However oscillatory modulation is a unique form of modulation; it shares neurotransmitters, and even synapses, with non-modulatory stimuli.
 As a result, homeostatic corrections for modulation must affect both the modulator and the driver \cite{Sherman1998}.
 Our work suggests this trade-off can be quite strong.
-To avoid this trade-off, sometimes anyway, we speculate that the long-time scale of evolution may favor different types of oscillation for different functional roles.
+To avoid this trade-off, sometimes anyway, we speculate that the long-time scale of evolution may sometimes favor different types of oscillation for different functional roles.
 
@@
 Even though both cortical and sub-cortical alpha generators are synaptically excitatory, this rhythm has however been associated with suppression of excitability \cite{Jensen2002,Bonnefond2012,Peterson2017}.
-While several competing explanations have been offered \cite{Bonnefond2012,Lange2013,Peterson2017} for this, our work raises another possibility that is complementary to the others.
+While several competing explanations have been offered for this \cite{Bonnefond2012,Lange2013,Peterson2017}, our work raises another possibility that is complementary to the others.
 Specifically, though we modelled hippocampal cells, the same Ca\textsuperscript{2+} homeostatic mechanisms exist in visual cortex (and the neocortex as a whole).
@@
 
-We do know, however, that oscillations are a ubiquitous feature of cortical activity, as is Ca\textsuperscript{2+}-mediated intrinsic homeostasis.
+We do know, however, that oscillations are a ubiquitous feature of cortical and sub-cortical activity, as is Ca\textsuperscript{2+}-mediated intrinsic homeostasis.
 This means the ingredients for oscillation and homeostasis to interact are omnipresent in both sub-cortical and cortical areas.
-This response might be very different than what we found.
-But that there could be broad range of effects makes our initial results in a simple model all the more intriguing.
+This response might be very different from what we found.
+But that there could be broad range of effects that make our initial results in a simple model all the more intriguing.
 
@@
 Meaning the qualitative properties of our model may also be present across phyla.
-That said the mechanisms for homeostatic interactions, along with their quantitative properties, depend on a number of factors specific to each cell and circuit.
+That said, the mechanisms for homeostatic interactions, along with their quantitative properties, depend on a number of factors specific to each cell and circuit.
 These include the duty cycle, power, and frequency of an oscillation, as well as on synaptic strengths and their location in the dendritic tree (an idea we return to below).
-It also depends on the other inputs into the cell, both from fast synaptic transmission and other (slower) modulators, as well a connection type; simulation studies suggest that recurrent connections can strongly interact with homeostatic regulation \cite{Harnack2015}.
+It also depends on the other inputs into the cell, both from fast synaptic transmission and other (slower) modulators, as well as connection type; simulation studies suggest that recurrent connections can strongly interact with homeostatic regulation \cite{Harnack2015}.
 The temporal and spatial scales of these factors will strongly influence Ca\textsuperscript{2+} dynamics, which is in turn central to governing what, if any, homeostatic effects oscillations may generate.
@@
 Synaptic homeostasis can also play key a role in combating oscillatory perturbations \cite{Cannon2016}.
-There is some evidence intrinsic homeostasis appear to be linked excitatory synaptic homeostasis \cite{Joseph2017}.
+There is some evidence that intrinsic homeostasis appears linked to excitatory synaptic homeostasis \cite{Joseph2017}.
 Still, the exact nature of the homeostatic response depends on how, and to what degree, oscillatory and other sensory or internally driven inputs share synapses.
@@
 \subsection*{Previous work}
-Homeostasis has been extensively studied in the rhythmic pacemaker present in the crab stomatogastric ganglion.
-Here homeostasis has been shown to stabilize self-organized oscillations \cite{Golowasch1999}, and interact with neuromodulation in a \textit{highly} state dependent way \cite{Marder2014,Marder2015,Marder2014}.
-Like our model though the oscillation present in the STG is tonic.
-This means it could in principle drive a second homeostatic response.
+Homeostasis has been extensively studied in the rhythmic pacemaker present in the crab and lobster stomatogastric ganglion, where homeostasis stabilizes these oscillations \cite{Golowasch1999}, and interacts with neuromodulation in a \textit{highly} state dependent way \cite{Marder2014,Marder2015,Marder2014}.
+Like our model though, the oscillation present in the stomatogastric ganglion is tonic.
+This means it could in principle drive a secondary homeostatic response.
 However, synapses in the stomatogastric ganglion circuit are predominately inhibitory \cite{Marder2015}.
-If our model predictions held up for this circuit we would expect there to be no second response.
+If our model predictions held up for this circuit we would expect there to be no secondary response.
 
 The interaction between homeostasis and oscillations has previously been considered in a simple model of cortex.
-But the oscillatory input here was treated as a signal, not a modulator.
 Cannon and Miller \citep{Cannon2017} explored how synaptic homeostasis can effectively minimize the effect of modulatory perturbations, thus maximizing mutual information between an incoming oscillatory signal and a single cell's firing pattern.
-Our analysis could be considered an inverse complement to \citep{Cannon2017}--we study how to minimize the perturbation caused by a modulatory oscillator, rather than how to maximize the transmission of an oscillator.
+But the oscillatory input in this work was treated as a signal, and not a modulator itself \citep{Cannon2017}.
+We believe this is an important distinction, and our analysis is an inverse complement to \citep{Cannon2017}.
+We studied how to minimize the perturbation caused by a oscillator, rather than how to maximize its transmission.
 
@@
  If our simple model is right, we conjecture that intrinsic homeostasis can explain why tonic theta rhythms in the hippocampus are synaptically inhibitory.
-To make this clear, consider the alternative.
-If the theta rhythm was strong, tonic, and synaptically excitatory, our model suggests this could lead to an equally strong--but opposing--homeostatic response.
+To make this clearer, consider the alternative: if the theta rhythm was strong, tonic, and synaptically excitatory, our model suggests this could lead to an equally strong--but opposing--homeostatic response.
 Such a response can mean that sodium conductance decreases, which in turn decreases the likelihood a hippocampal neuron could respond to any given stimulation.
@@
 This may also explain why neocortical oscillations tend to be short and bursty, and why some neurological disorders, such as Parkinson's disease, are associated with prolonged excitatory rhythms.
-
-% Computational modeling plays a critical role in neuroscience.
-Explicit computational modeling moves the field from "word model" theories to self-consistent mathematical formalizations [REF].
-While these models can be wrong, they offer testable novel venues for research that are often difficult to find due to the immense complexity of the rich, dynamic neuronal environment.
 
@@
 The synaptic weights for all $N \rightarrow M$ connections $w$ were independently sampled from a uniform distribution, $w \sim \mathcal{U}(5, 50)$ $\mu$S.
-The firing rate of the oscillating population was governed by sinusoidal pacemaker, with amplitude $A$ and frequency $f$, with the exact form $\frac{r_o}{2} (1 + \text{sin}(2 \pi f t))$.
+The firing rate of the oscillating population was governed by a biased sinusoidal pacemaker, with amplitude $\frac{r_o}$ and frequency $f$, with the exact form $\frac{r_o}{2} (1 + \text{sin}(2 \pi f t)) + r_b$.
 As a result, $r_o$ defines the peak firing of the oscillation.
@@
 
-Overall membrane dynamics were governed by these internal ion conductances, a variable bias current $I_\text{bias}$, and the excitatory synaptic input term $I_S$, which contains both stimulus and oscillatory terms.
-All synaptic input was in turn governed a single exponential kinetics, which we denote generically using an $x$ subscript below.
-Though each synaptic input had different inputs, all models shared the same parameters.
-That is, $\tau_x = \tau_b = \tau_s = \tau_o$ and $\bar{g_x} = \bar{g_b} = \bar{g_s} = \bar{g_o}$.
+Overall membrane dynamics were governed by these internal ion conductances, and two synaptic input terms, $I_s$ and $I_o$.
+These are, respectively, the stimulus and oscillatory inputs.
+All synaptic input, both stimulus and oscillation were governed by single exponential kinetics, which we denote generically using an $x$ subscript below.
 
 \begin{equation}
-    I_S = -g_s (V_e - V) + -g_o (V_o - V) 
+    \label{eq:I_x1}
+    I_{\text{x}} = -g_x (V_x - V)
 \end{equation}
 \begin{equation}
+    \label{eq:I_x2}
     \tau_x \frac{dg_x}{dt} = -g_x + \bar{g_x} \delta(t - t_x^j)
@@
 
+As a concrete example of using the abstract form in Eq~\ref{eq:I_x1}-\ref{eq:I_x2}, let's define the stimulus.
+The stimulus was always excitatory, so to define $I_s$ we substitute in AMPAergic voltage, conductance, and time constant values taken from Table~\ref{table:params} in for $x$ (above), giving Eqs.~\ref{eq:I_s1}-\ref{eq:I_s2}.
+
+\begin{equation}
+    \label{eq:I_s1}
+    I_{s} = -g_e (V_e - V)
+\end{equation}
+\begin{equation}
+    \label{eq:I_s2}
+    \tau_e \frac{dg_e}{dt} = -g_e + \bar{g_e} \delta(t - t_s^j)
+\end{equation}
+
+To define $I_o$ the model makes a similar set of substitutions, choosing either AMPAergic or GABAergic values, depending on the which kind of oscillation we wish to simulate.
+
+Having defined both internal dynamics and synaptic dynamics we arrive at final model, stated in terms of its biophysical currents (Eq.~\ref{eq:HHH}).
+
 \begin{equation}
 \label{eq:HHH}
-C \frac{dV}{dt} = I_l + I_{Na} + I_{K} + I_{KCa} + I_{Ca} + I_{\text{bias}} + I_{S} 
+C \frac{dV}{dt} = I_l + I_{Na} + I_{K} + I_{KCa} + I_{Ca} + I_{o} + I_{s} 
 \end{equation}
 
-The intrinsic excitability is regulated by altering both inward and outward conductances in response to changes in Ca\textsuperscript{2+} concentration.
-Following the previous work \cite{LeMasson1993} and \cite{Siegel1994}, we modeled this by allowing the maximal conductances $\bar{g_\text{Na}}$, $\bar{g_\text{K}}$, $\bar{g_\text{KCa}}$, and $\bar{g_\text{Ca}}$ to non-linearly vary in response to changes in Ca\textsuperscript{2+} concentration, $\text{Ca}$.
-During homeostatic equilibration, conductances drifted until the target Ca\textsuperscript{2+} concentration was met, denoted as $C_T$.
-In a control experiment a range of $C_T$ values were explored (Figure \ref{fig:f5}), though simulations default to 0.03 mM; the value the system reaches with stimulation ($r_s = 6$ Hz) without modulation ($r_o = 0$).
-The $\pm$ symbol in equation \ref{eq:dgdt} denotes the direction of ion flow and is $(+)$ for inward going currents (Na and Ca) and $(-)$ for outward going Potassium.
+% TODO add noise term here. Replaces r_b. You have not mentioned this until now why you idiot?
+
+Eq~\ref{eq:HHH} would mark the end of a traditional Hodgkin-Huxley definition.
+However modelling intrinsic homeostasis means altering both inward and outward conductances in response to changes in Ca\textsuperscript{2+} concentration.
+Following the previous work of \cite{LeMasson1993} and \cite{Siegel1994}, we modeled this by having maximal conductances $\bar{g_\text{Na}}$, $\bar{g_\text{K}}$, $\bar{g_\text{KCa}}$, and $\bar{g_\text{Ca}}$ non-linearly vary with the Ca\textsuperscript{2+} concentration, $\text{Ca}$ (Eq.\ref{eq:dgdt}).
+In this scheme and during homeostatic equilibration, conductances drift until the target Ca\textsuperscript{2+} concentration is met, denoted as $C_T$.
+In a control experiment a range of $C_T$ values were explored (Figure \ref{fig:f5}), though simulations default to 0.003 mM; the value the system reaches with stimulation ($r_s = 6$ Hz) without modulation ($r_o = 0$).
+The $\pm$ symbol in equation \ref{eq:dgdt} denotes the direction of ion flow and is $(+)$ for inward going currents and $(-)$ for outward going.
 
@@
 $r_s$ & 6 (Hz) & Stimulus firing rate \\
+$r_b$ & 2 (Hz) & The ``background'' firing rate \\
 $C$ & 1 ($\mu$F) & Membrane capacitance \\

### 2020-05-04 main.tex
Reviewer response 2 - complete

@@
 We modulate this population with neural oscillations.
-Oscillatory input for each cell in Hodgkin-Huxley population was modelled as a set of about $M=100$ Poisson neurons (Figure~\ref{fig:f1}\textbf{b}-\textbf{c}).
+Oscillatory input for each cell in Hodgkin-Huxley population was modeled as a set of about $M=100$ Poisson neurons (Figure~\ref{fig:f1}\textbf{b}-\textbf{c}).
 With a probability of 0.1 these were randomly drawn from a pool of 1000 Poisson neurons, each of which was independently simulated.
-To change the level of oscillatory modulation strength, we alter the firing rate of the entire Poisson population.
-Average synaptic strengths were fixed, but drawn from a uniform distribution for each simulation (see \textit{Methods}).
+To change the level of oscillatory modulation strength, we altered the firing rate of the entire Poisson population.
+Average synaptic strengths were fixed, but drawn from a uniform distribution for each s
+
+imulation (see \textit{Methods}).
 An oscillation's temporal duration was simulated as either tonic (Figure~\ref{fig:f1}\textbf{b}, top), which lasted for the entire experiment (20 sec), or as a 3-cycle burst (Figure~\ref{fig:f1}\textbf{b}, bottom).
@@
 While our simple model has been shown by others to recapitulate hippocampal pyramidal cell's activity in laboratory settings \cite{LeMasson1993}, it is a \textit{very simple} model, and validating our work does require new experimental testing.
-This could be done, for example, by using prolonged excitatory gamma oscillations--arising from direct stimulation of either the dentate gyrus or select areas in CA3 \cite{Csicsvari2003}--to drive a possible homeostatic response in CA1.
+This could be done, for example, by using prolonged excitatory gamma oscillations--arising from direct stimulation of either the medial entorhinal cortex or CA3--to drive a possible homeostatic response in CA1 \cite{Csicsvari2003,Colgin2009}.
 
@@
 The stimulus window was matched to the length of time 3 cycles of oscillation would take.
+To generate this curve the model was equilibrated for 20 seconds, under a 1.5 Hz firing rate and 8 Hz frequency rhythm, after which the current impulse that we described above was applied.
+During FI-curve estimation these was no oscillation present.
 }
@@
 Our burst/tonic distinction is therefore a false (but useful) dichotomy.
-To find a better estimate of how long an oscillation would need to last, on average, to generate a homeostatic response we ran some long simulations using a more realistic homeostatic time constant ($\tau_h$ = 600 seconds; up from 20 seconds).
+To find a better estimate of how long an oscillation would need to last, on average, to generate a homeostatic response we ran some long simulations using a more realistic homeostatic time constant ($\tau_h$ = 600 seconds; up from 4 seconds).
 It seems an oscillation would need to be present \textit{at least} 10\% of the time to generate a response (Figure~\ref{fig:f5}\textbf{c}).
-Real theta rhythms drift in frequency \cite{Buzsaki2015}.
+Jacobs \cite{Jacobs2014} estimated the prevalence of hippocampal theta to be about 40\% of the time in rats, and 10\% in humans.
+So even for humans, who have slower and less stable theta rhythm(s) \cite{Goyal2018}, we'd expect an homeostatic response from AMPAergic theta, if it existed.
+
+Real theta rhythms drift in frequency \cite{Buzsaki2015,Jacobs2014}.
 To test if small differences in frequency have a notable effect, we simulated small drifts in frequency of theta ranging from 6-10 Hz.
@@
 Put another way: can homeostasis explain why some oscillations, such as hippocampal theta, tend to appear as tonic rhythms while other oscillations (like those in cortex) tend to appear as bursts?
+\cite{Peterson2017,Cole2018,vanEde2018a}
 
-Oscillations are generally conceived to play a functional role in modulating spiking {Cite Varela}{Cite Fries}, by grouping action potentials and improving signal to noise \cite{Chen2013,Zhou2015,Voytek2015a,Peterson2017}.
+Oscillations are generally conceived to play a functional role in modulating spiking \cite{Fries2015a,vanEde2018b} by grouping action potentials and improving signal to noise \cite{Chen2013,Zhou2015,Voytek2015a,Peterson2017}.
 Absent the effects of homeostasis, our model clearly demonstrates this: with increasing modulation strength, excitatory oscillations show increased firing rate and synchrony.
@@
 One striking example is in motor cortical regions, where beta (~12-30 Hz) bursts are prevalent and likely functional.
-Specifically, beta bursts are very short--sometimes lasting only one or two cycles \cite{Sherman2016}--and relate to self-timed movements \cite{Feingold2015}.
+Specifically, beta bursts are very short--sometimes lasting only one or two cycles \cite{Sherman2016,Cole2019,Jones2016}--and relate to self-timed movements \cite{Feingold2015}.
 Patients with Parkinson's disease show increased motor rigidity and bradykinesia, symptoms associated with prolongation of beta bursts \cite{Tinkhauser2017}.
@@
 While several competing explanations have been offered for this \cite{Bonnefond2012,Lange2013,Peterson2017}, our work raises another possibility that is complementary to the others.
-Specifically, though we modelled hippocampal cells, the same Ca\textsuperscript{2+} homeostatic mechanisms exist in visual cortex (and the neocortex as a whole).
+Specifically, though we modeled hippocampal cells, the same Ca\textsuperscript{2+} homeostatic mechanisms exist in visual cortex (and the neocortex as a whole).
 This means that strong and tonic alpha oscillations, combined with homeostasis, should directly suppress population firing in visual cortex.
@@
 
-% TODO add noise term here. Replaces r_b. You have not mentioned this until now why you idiot?
-
 Eq~\ref{eq:HHH} would mark the end of a traditional Hodgkin-Huxley definition.

### 2020-05-04 main.tex
Fix - math typo

@@
 The synaptic weights for all $N \rightarrow M$ connections $w$ were independently sampled from a uniform distribution, $w \sim \mathcal{U}(5, 50)$ $\mu$S.
-The firing rate of the oscillating population was governed by a biased sinusoidal pacemaker, with amplitude $\frac{r_o}$ and frequency $f$, with the exact form $\frac{r_o}{2} (1 + \text{sin}(2 \pi f t)) + r_b$.
+The firing rate of the oscillating population was governed by a biased sinusoidal pacemaker, with amplitude $\frac{r_o}{2}$ and frequency $f$, with the exact form $\frac{r_o}{2} (1 + \text{sin}(2 \pi f t)) + r_b$.
 As a result, $r_o$ defines the peak firing of the oscillation.
